### 2025-03-31 Paper7.tex
Update local file

\documentclass[preprint,12pt]{elsarticle}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{array}
\usepackage{graphicx}
\usepackage{placeins}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{subcaption}
\usepackage{mwe}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{adjustbox}
\usepackage{multirow}
\usepackage{hyperref}
\usepackage[noabbrev,nameinlink]{cleveref}
\usepackage[title]{appendix}
\usepackage[utf8]{inputenc}
\usepackage{stanli}
\usepackage{tikz}
\usepackage{siunitx}
\usepackage{xspace}
\usepackage{lmodern}
\usepackage{nomencl}
\makenomenclature
\usepackage{longtable}
\usepackage{lineno}
\usetikzlibrary{arrows.meta} % Load the arrows.meta library
\usetikzlibrary{shapes,positioning}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\AtBeginEnvironment{appendices}{\crefalias{section}{appendix}}

% Derivatives should be upright d (JSV requires it)
% I don't have the docs for the derivate package with me, so I'll just do this bad hack for now.
% Could update all the ODEs to use the derivative package
% Update: hack didn't work due to \d already defined somewhere - leaving this for now
%\DeclareMathOperator{\d}{\textnormal{d}}

\newcommand{\kph}{~\unit{\kilo\metre\per\hour}\xspace}
\newcommand{\um}{~\unit{\metre}\xspace}
\newcommand{\ukg}{~\unit{\kilogram}\xspace}
\newcommand{\urho}{~\unit{\kilogram\per\meter}\xspace}
\newcommand{\uI}{~\unit{\kilogram\square\metre}\xspace}
\newcommand{\urd}{~\unit{\newton\square\metre}\xspace}
\newcommand{\ustif}{~\unit{\newton\per\metre}\xspace}
\newcommand{\udamp}{~\unit{\newton\second\per\metre}\xspace}
\newcommand{\DAF}{\textnormal{DAF}}
% Hacky - is there a better upright i?
%\newcommand{\ii}{\textnormal{i}}
\newcommand{\id}{\mathrm{d}\mkern1mu}
\newcommand{\ii}{\mathrm{i}\mkern1mu}
\newcommand{\nukg}{\unit{\kilogram}\xspace}
\newcommand{\nustif}{\unit{\newton\per\metre}\xspace}
\newcommand{\nudamp}{\unit{\newton\second\per\metre}\xspace}
\newcommand{\nugpa}{\si{\giga\pascal}\xspace}
\begin{document}

% I made some comments about x/L being xi - it's not used much so you can ignore these.
% There are some cases where I added text in [square brackets] and not necessarily in a latex comment
% Check the bib file - two references had no journal - I added these as XXXX
% Section 4 does too much - could you consider splitting this section up a bit? 
% I did add some more subsubsections to try structure section 4 a bit better, but I'm not sure it's optimal yet.
\begin{frontmatter}
\title{The dynamic stiffness matrix based on the extended separation-of-variables type solutions for the free vibration of orthotropic rectangular thin plates}
\author[1]{Shiyi Mei}\ead{shiyi.mei1@monash.edu}
\author[1]{Colin Caprani\corref{cor1}}\ead{colin.caprani@monash.edu}
\author[2]{Daniel Cantero}\ead{daniel.cantero@ntnu.no}
\cortext[cor1]{Corresponding author}
\affiliation[1]{organization={Department of Civil Engineering, Monash University},city={Melbourne},state={Victoria},country={Australia}}
\affiliation[2]{organization={Department of Structural Engineering, Norwegian University of Science \& Technology NTNU},city={Trondheim}, country={Norway}}
\begin{abstract}
The dynamic stiffness matrix based on the extended separation-of-variables mode functions is developed for the free vibration analysis of an orthotropic rectangular thin plate with general homogeneous boundary conditions.  
The governing differential equation and boundary conditions are derived from Rayleigh’s principle.  
Based on the boundary conditions, the dynamic stiffness matrix is formulated and solved using an improved Wittrick-Williams algorithm.  
In this improved algorithm, simply supported frequencies are required instead of fully clamped frequencies, and the closed-form expression for the simply supported frequencies is provided, enhancing the efficiency and systematicity of solving the eigenvalue problem.  
The proposed method is validated by the numerical experiments.
\end{abstract}
\end{frontmatter}
\linenumbers
\section{Introduction}
Rectangular plates play an important role in various engineering fields, including civil, mechanical, and aerospace engineering \citep{biancolini2005approximate}. 
The free vibration of plates has been a fundamental research problem for over two centuries. 
The earliest exact solutions for this problem are the Navier \citep{navier1823extrait} and Levy \citep{levy1899equilibre} solutions, which require at least one pair of opposite edges to be simply supported or guided.
To solve problems with other boundary conditions, approximate solutions such as the Rayleigh–Ritz method \cite{leissa1973free} and the Galerkin method \cite{laura1967study} have been widely applied. 
For these approximation methods, beam functions, polynomials, trigonometric functions, and their combinations \cite{li2004vibration} are commonly used as the assumed approximate functions.
The accuracy of these solutions depends on how well the assumed approximate functions represent the displacement of the plate.

Besides the approximation methods, several analytical methods have been developed over the past decades, including the Kantorovich-Krylov method \cite{kantorovich1958approximate,kerr1968extension}, the symplectic eigenfunction expansion method \cite{zhong1995new,xing2009new2}, the separation-of-variable (SOV) method \cite{xing2009new}, the dynamic stiffness matrix (DSM) method \cite{banerjee1997dynamic}, and series expansion-based methods \cite{xing2022review}.
The series expansion-based methods include the superposition method \cite{timoshenko1940theory,gorman2005free}, Fourier series method \cite{khov2009accurate,li2009exact}, the finite integral transform method \cite{li2009finite,zhong2013free}, and other series methods. 
These methods represent the plate displacement in terms of an infinite series and mostly are capable of handling any general boundary conditions. 
However, sufficient truncation of the series is required to ensure the accuracy and convergence of the results, and the eigenvalue equation is generally difficult to express explicitly.
Therefore, solving the corresponding eigenvalue problem can be computationally expensive.

Despite being a powerful method for the dynamic analysis of plate assemblies, the finite element method (FEM) requires a sufficient number of elements and is computationally expensive to accurately capture higher-order modes.
Thus, the DSM method is developed as an accurate and efficient analytical approach to alternatively solve complex plate structures \cite{boscolo2011dynamic,fazzolari2013exact}. 
The DSM can be considered as an analytical FEM since the mode functions of the plate are expressed by analytical solutions, where Levy-type solution \cite{ghorbel2015dynamic} or components of infinite Fourier series \cite{banerjee2015dynamic,liu2016free} are applied.
To avoid solving the cumbersome transcendental frequency equation directly, the Wittrick-Williams (W-W) algorithm \cite{wittrick1971general} is applied to the eigenvalue problem.
The W-W algorithm determines the lower and upper bounds of natural frequencies rather than solving the frequency equation directly. 
Thus, the DSM has the potential to be effectively and systematically solved using the W-W algorithm.
However, a critical part in applying the W-W algorithm is to priorly determine all natural frequencies of the fully clamped structure within the interested frequency range.
Strategies such as using a sufficiently fine mesh or including a sufficient number of terms in series expansions \cite{banerjee2015dynamic} can ensure that all fully clamped frequencies are accounted for, thereby maintaining the accuracy of the algorithm. However, these approaches are computationally expensive and complex, posing a significant obstacle to the wider adoption and application of the DSM method based on the W-W algorithm \cite{han2018improved}.
To resolve the fully clamped plate problem, \citet{liu2015exact} suggested that the fully clamped frequencies can be indirectly obtained from the simply supported plate problem, where the Navier solution serves as the analytical solution. This provides a significant enhancement to the W-W algorithm, increasing the efficiency of applying DSM methods. 
However, since infinite-term series expansion-based solutions are used for plate analysis instead of explicit and closed-form solutions, a sufficient number of truncation terms is required to ensure accuracy and capture the necessary modal degrees.

Inspired by the Navier and Levy solutions, \citet{xing2009new} proposed the separation-of-variables (SOV) method, which provides concise and explicit eigensolutions.
The mode shape function has a separable form, $\phi(x) \psi(y)$, requiring only one $\phi(x)$ and one $\psi(y)$ for each mode order, allowing each eigenvalue equation to be explicitly expressed.
However, this SOV method is not suitable to deal with plate with free boundary conditions.
Therefore, an extended SOV method \cite{xing2018overall, xing2020extended} based on the Rayleigh quotient is proposed to accommodate plates with all classical boundary conditions, i.e., simply supported, clamped, guided, and free.
Based on the Rayleigh quotient model, alternative iterative and improved SOV methods have been subsequently proposed \cite{xing2020improved}.
Although SOV methods provide concise closed-form analytical solutions, solving the highly nonlinear eigenvalue equations is required.

In this study, the free vibration of orthotropic plates with general boundary conditions is analyzed by extending the SOV method based on the Rayleigh quotient to obtain the solution for elastically restrained plates.
Subsequently, dynamic stiffness matrices in both the \(x\) and \(y\) directions, derived from the SOV-form solutions, are developed to solve the eigenvalue problem.
By applying the 'hypothetical structure method', a modified W-W algorithm is developed, requiring only the natural frequencies of the simply supported structure instead of the necessity of fully clamped frequencies.
The concise closed-form expression for the simply supported frequencies is directly obtained from the eigenvalue equation based on the SOV method, eliminating the extensive computational effort typically required in DSM methods to determine all fully clamped frequencies within the frequency range.  
Therefore, this approach can improve the systematic and effective application of the DSM method depending on the W-W algorithm.
Alternatively, it can also be considered as an efficient technique for solving transcendental eigenvalue equations in the SOV method.

\FloatBarrier
\section{Mathematical model}\label{sec:Mathematical model}
\begin{figure}[!htbp]
	\centering
	\resizebox{0.8\textwidth}{!}
	{
		\input{figs/general_plate_tikz.tex}
	}
	\caption{\small The orthotropic rectangular plate with all edges elastically restrained.} 
	\label{fig:platemode}
\end{figure}
Consider a thin orthotropic rectangular plate of length \(2a\) and width \(2b\), with all four edges restrained by vertical translational springs \(k^v\) and rotational springs \(k^r\), as shown in \Cref{fig:platemode}. 
The coordinate origin is located at the center of the plate.

The governing differential equation for the free vibration of a thin orthotropic plate is given by \citep{xing2020improved}:  
%
\begin{equation}\label{eq:governing_EOM}
	D_{11}\frac{\partial^4w}{\partial \xi^4} + 2D_3\alpha^2\frac{\partial^4w}{\partial \xi^2 \partial \eta^2} + D_{22}\alpha^4\frac{\partial^4w}{\partial \eta^4} = \rho h\alpha^4\omega^2w,
\end{equation}
%
where the bending stiffness parameters are defined as:
%
\begin{equation}\label{eq:bd_stiff}
	\begin{split}
		&D_{11} = \frac{E_1h^3}{12(1-v_{12}v_{21})}, \quad D_{22} = \frac{E_2h^3}{12(1-v_{12}v_{21})}, \\  
		&D_{66} = \frac{G_{12}h^3}{12}, \quad D_{12} = v_{12}D_{22} = v_{21}D_{11}, \quad D_3 = D_{12} + 2D_{66},
	\end{split}
\end{equation}
%
where \(\rho\) and \(h\) denote the mass density and thickness of the plate, respectively; \(E_1\) and \(E_2\) are the Young’s moduli in the \(x\)- and \(y\)-directions, respectively; \(G_{12}\) is the shear modulus, and \(v_{12}\) and \(v_{21}\) are the Poisson’s ratios.

Instead of solving the free vibration of the thin orthotropic plate using \Cref{eq:governing_EOM}, it is suggested that the vibration of the thin plate can also be solved using the Rayleigh quotient variational principle \cite{xing2018overall}:
%
\begin{equation}\label{eq:Rayleigh}
	\begin{split}
		\delta U_{mag} &= \omega^2\,\delta T_0,
	\end{split}
\end{equation}
%
where \(\delta\) denotes variation, \(U_{mag}\) is the magnitude of the potential energy of the plate, and \(\omega^2 T_0\) represents the magnitude of the kinetic energy of the plate.
The potential energy of the plate can be expressed as \cite{xing2020extended}:
%
\begin{equation}\label{eq:poten_energy}
	\begin{split}
		U^{I} &= \frac{1}{2}\iint \Big[D_{11}\left(\frac{\partial^2 W}{\partial x^2}\right)^2 + 2D_{12}\frac{\partial^2 W}{\partial x^2}\frac{\partial^2 W}{\partial y^2} + D_{22}\left(\frac{\partial^2 W}{\partial y^2}\right)^2 \\
		&\qquad + 4D_{66}\left(\frac{\partial^2 W}{\partial x \partial y}\right)^2\Big] \, \mathrm{d}x \, \mathrm{d}y.
	\end{split}
\end{equation}
%
And the kinetic energy is:
%
\begin{equation}\label{eq:kine_energy}
	\begin{split}
		T &= \frac{1}{2}\iint \rho h \left(\frac{\partial W}{\partial t}\right)^2 \, \mathrm{d}x \, \mathrm{d}y.
	\end{split}
\end{equation}

Assuming the solution of the deflection \( W(x, y; t) = w(x, y) e^{i \omega t} \) for harmonic plate motion, 
where \( \ii = \sqrt{-1} \), \( w(x, y) \) is the mode shape, and \( \omega \) is the radial frequency.  
By substituting \( W(x, y; t) = w(x, y) e^{i \omega t} \) into \Cref{eq:poten_energy,eq:kine_energy} and expressing the system in dimensionless coordinates, we have:
%
\begin{equation}\label{eq:poten_energy_mag}
	\begin{split}
		U_{\text{mag}}^{I} &= \frac{ab}{2} \iint \left[ \frac{D_{11}}{a^4} \left( \frac{\partial^2 w}{\partial \xi^2} \right)^2 + \frac{2 D_{12}}{a^2 b^2} \frac{\partial^2 w}{\partial \xi^2} \frac{\partial^2 w}{\partial \eta^2} + \frac{D_{22}}{b^4} \left( \frac{\partial^2 w}{\partial \eta^2} \right)^2 \right. \\
		&\quad \left. + \frac{4 D_{66}}{a^2 b^2} \left( \frac{\partial^2 w}{\partial \xi \partial \eta} \right)^2 \right] \, \mathrm{d} \xi \, \mathrm{d} \eta,
	\end{split}
\end{equation}
%
and
%
\begin{equation}\label{eq:kine_energy_mag}
	\begin{split}
		T &= \omega^2 \frac{ab}{2} \rho h \iint w^2 \, \mathrm{d} \xi \, \mathrm{d} \eta = \omega^2 T_0
	\end{split},
\end{equation}
%
where \( \alpha = a / b \) is the aspect ratio; \( \xi = x / a \) and \( \eta = y / b \) are the normalized coordinates.
The separable form of the mode shape function \( w(\xi, \eta) \) is given by:
%
\begin{equation}\label{eq:displace_func}
	w(\xi, \eta) = \phi(\xi) \psi(\eta),
\end{equation}
%
where \( \phi(\xi) \) and \( \psi(\eta) \) can be expressed as:
\begin{subequations}\label{eq:phi_psi_solution}
	\begin{align}
		\phi(\xi) &= A_1 \sin{(\alpha_1 \xi)} + A_2 \cos{(\alpha_1 \xi)} + A_3 \sinh{(\beta_1 \xi)} + A_4 \cosh{(\beta_1 \xi)},\label{eq:phi_solution1}\\
		\psi(\eta) &= B_1 \sin{(\alpha_2 \eta)} + B_2 \cos{(\alpha_2 \eta)} + B_3 \sinh{(\beta_2 \eta)} + B_4 \cosh{(\beta_2 \eta)}. \label{eq:phi_solution2}
	\end{align}
\end{subequations}

It should be noted that \Cref{eq:governing_EOM} represents the strong-form governing equation of the orthotropic plate, while \Cref{eq:Rayleigh} is the weak-form governing equation, with the latter being equivalent to the former.
Based on \Cref{eq:Rayleigh}, the frequencies $\omega_x$ and $\omega_y$, corresponding to the mode shapes $\phi(\xi)$ and $\psi(\eta)$, respectively, are assumed to be independent of each other.

\subsection{Dynamic stiffness matrix corresponding to $\omega_x$}\label{sec:DSMx}
For given general homogeneous boundary conditions, we can first assume that the mode shape $\psi(\eta)$ corresponding to the $y$ direction is known. 
Supposing the edges of the plate in both the \(x\)- and \(y\)-directions are elastically restrained by homogeneous vertical translational and rotational springs.
The vertical translational and rotational springs at the \(\xi = -1\) end are defined as \(k^v_{\xi = -1}\) and \(k^r_{\xi = -1}\), respectively, and at the \(\xi = 1\) end as \(k^v_{\xi = 1}\) and \(k^r_{\xi = 1}\), respectively. 
Thus, the potential energy along the supported edge in the \(x\)-direction can be expressed by:
%
\begin{equation}\label{eq:poten_energy_edgex}
	\begin{split}
		U^{II} &=  \int\left[k^r_{\xi=-1} \left( \frac{\partial W}{\partial x} \right)^2 + k^v_{\xi=-1} \left( W \right)^2\right]_{x=-a} \id y\\
		&\quad +  \int\left[k^r_{\xi=1} \left( \frac{\partial W}{\partial x} \right)^2 + k^v_{\xi=1} \left( W \right)^2\right]_{x=a} \id y.
	\end{split}
\end{equation}
%
From \Cref{eq:poten_energy_edgex}, the magnitude of total potential energy along the edges in the \(x\)-direction is obtained as:
%
\begin{equation}\label{eq:poten_energy_edgex2}
	\begin{split}
		U^{II}_{mag} &= ab \int\left[\frac{k^r_{\xi=-1}}{a^3} \left( \frac{\partial w}{\partial \xi} \right)^2 + \frac{k^v_{\xi=-1}}{a} \left( w \right)^2\right]_{\xi=-1} \, \id \eta \\
		&\quad + ab \int\left[\frac{k^r_{\xi=1}}{a^3} \left( \frac{\partial w}{\partial \xi} \right)^2 + \frac{k^v_{\xi=1}}{a} \left( w \right)^2\right]_{\xi=1} \, \id \eta.
	\end{split}
\end{equation}
%
The magnitude of potential energy of the plate in the \(x\)-direction can be obtained from \Cref{eq:poten_energy_mag,eq:poten_energy_edgex2} as:
%
\begin{equation}\label{eq:poten_energy_tolx}
	\begin{split}
		U_{mag} &= U^{I}_{mag} + U^{II}_{mag} \\
		&= \frac{ab}{2} \iint \left[ \frac{D_{11}}{a^4} \left( \frac{\partial^2 w}{\partial \xi^2} \right)^2 + \frac{2 D_{12}}{a^2 b^2} \frac{\partial^2 w}{\partial \xi^2} \frac{\partial^2 w}{\partial \eta^2} + \frac{D_{22}}{b^4} \left( \frac{\partial^2 w}{\partial \eta^2} \right)^2 \right. \\
		&\quad \left.  + \frac{4 D_{66}}{a^2 b^2} \left( \frac{\partial^2 w}{\partial \xi \partial \eta} \right)^2 \right] \, \id\xi \, \id\eta + ab \int \left[\frac{k^r_{\xi=1}}{a^3} \left( \frac{\partial w}{\partial \xi} \right)^2 + \frac{k^v_{\xi=1}}{a} \left( w \right)^2 \right]_{\xi=1} \, \id \eta \\
		&\quad  + ab \int \left[\frac{k^r_{\xi=-1}}{a^3} \left( \frac{\partial w}{\partial \xi} \right)^2 + \frac{k^v_{\xi=-1}}{a} \left( w \right)^2 \right]_{\xi=-1} \, \id \eta
	\end{split}
\end{equation}
%
By substituting \Cref{eq:displace_func} into \Cref{eq:poten_energy_tolx}, we have:
%
\begin{equation}\label{eq:poten_energy_tolx2}
	\begin{split}
		U_{mag}&=U^{I}_{mag}+U^{II}_{mag}\\
		&=\frac{ab}{2} \int_{-1}^{1} \left[ \frac{D_{11}}{a^4}S_1 \left( \frac{\id^2\phi }{\id \xi^2} \right)^2 + \frac{2 D_{12}}{a^2 b^2}S_2 \frac{\id^2\phi }{\id \xi^2}\phi  + \frac{D_{22}}{b^4}S_4  \phi^2 \right. \\
		&\quad \left. + \frac{4 D_{66}}{a^2 b^2}S_3 \left( \frac{\id \phi }{\id \xi}\right)^2 \right] \, \id\xi +abS_1 \left[\frac{k^r_{\xi=-1}}{a^3} \left( \frac{\id \phi}{\id \xi} \right)^2+ \frac{k^v_{\xi=-1}}{a} \left( \phi \right)^2\right]_{\xi=-1}\\
		&\quad+abS_1 \left[\frac{k^r_{\xi=1}}{a^3} \left( \frac{\id \phi}{\id \xi} \right)^2+ \frac{k^v_{\xi=1}}{a} \left( \phi \right)^2\right]_{\xi=1},
	\end{split}
\end{equation}
%
where the integral parameters are defined as:
%
\begin{equation}\label{eq:inte_phi_text}  
	\begin{split}  
		&S_1 = \int_{-1}^{1} \psi^2 \, \id\eta, \\  
		&S_2 = \int_{-1}^{1} \left( \frac{\id^2 \psi}{\id \eta^2} \psi \right) \, \id\eta, \\  
		&S_3 = \int_{-1}^{1} \left( \frac{\id \psi}{\id \eta} \right)^2 \, \id\eta, \\  
		&S_4 = \int_{-1}^{1} \left( \frac{\id^2 \psi}{\id \eta^2} \right)^2 \, \id\eta.  
	\end{split}  
\end{equation}   
%
By taking \Cref{eq:displace_func} into account, the coefficient \(T_0\) of the kinetic energy in \Cref{eq:kine_energy_mag} for the plate can be expressed as:
%
\begin{equation}\label{eq:kine_energyx}
	\begin{split}
		T_0 = \frac{ab}{2}\rho h\iint w^2 \, \id \xi \, \id \eta = \frac{ab}{2}\rho h S_1\int_{-1}^{1} \phi^2 \, \id \xi.
	\end{split}
\end{equation}

Take the Rayleigh principle in the form:
%
\begin{equation}\label{eq:Rayleighx}
	\begin{split}
		\delta U_{mag} &= \omega_x^2 \, \delta T_0.
	\end{split}
\end{equation}
%
By substituting \Cref{eq:poten_energy_tolx2,eq:kine_energyx} into \Cref{eq:Rayleighx}, relieve \(\delta \phi\) and \(\delta\frac{\id  \phi}{\id \xi}\) in \Cref{eq:Rayleighx} by variation calculus, yielding:
%
\begin{equation}\label{eq:Rayleighx2}
	\begin{split}
		0 &= \int_{-1}^{1} \bigg[ \frac{D_{11}}{a^4} S_1 \frac{\id^4 \phi}{\id \xi^4} 
		+ \bigg( \frac{2D_{12}}{a^2b^2} S_2 - \frac{4D_{66}}{a^2b^2} S_3 \bigg) \frac{\id^2 \phi}{\id \xi^2} \\
		&\quad + \bigg( \frac{D_{22}}{b^4} S_4 - \omega_x^2 \rho h S_1 \bigg) \phi \bigg] \delta \phi \, \id \xi \\
		&\quad + \frac{2k^v_{\xi=-1}}{a} S_1 \big(\phi \delta \phi\big)_{\xi=-1} 
		+ \frac{2k^v_{\xi=1}}{a} S_1 \big(\phi \delta \phi\big)_{\xi=1} \\
		&\quad + \bigg[ \bigg( \frac{4D_{66}}{a^2b^2} S_3 - \frac{D_{12}}{a^2b^2} S_2 \bigg) \frac{\id \phi}{\id \xi} 
		- \frac{D_{11}}{a^4} S_1 \frac{\id^3 \phi}{\id \xi^3} \bigg] \delta \phi \Big|^{\xi=1}_{\xi=-1} \\
		&\quad + \bigg( \frac{D_{12}}{a^2b^2} S_2 \phi + \frac{D_{11}}{a^4} S_1 \frac{\id^2 \phi}{\id \xi^2} \bigg) 
		\delta\frac{\id  \phi}{\id \xi} \Big|^{\xi=1}_{\xi=-1} \\
		&\quad + \frac{2k^r_{\xi=-1}}{a^3} S_1 \bigg(\frac{\id \phi}{\id \xi}\delta \frac{\id \phi}{\id \xi}\bigg)_{\xi=-1} 
		+ \frac{2k^r_{\xi=1}}{a^3} S_1 \bigg(\frac{\id \phi}{\id \xi} \delta\frac{\id \phi}{\id \xi}\bigg)_{\xi=1}.
	\end{split}
\end{equation}
%
Thus, the governing differential equation in the \( x \)-direction can be obtained from the integration part in \Cref{eq:Rayleighx2}:
%
\begin{equation}\label{eq:ODEx}
		\frac{\id^4 \phi}{\id \xi^4}+ 2\alpha^2\left(\frac{D_{12}S_2}{D_{11}S_1}-2\frac{D_{66}S_3}{D_{11}S_1}\right)\frac{\id^2 \phi}{\id \xi^2}+\left(\alpha^4\frac{D_{22}S_4}{D_{11}S_1}-a^4\Omega_x^4\right)\phi=0,
\end{equation}
%
where \(\Omega_x = \sqrt[4]{\omega_x^2 \rho h / D_{11}}\).
By substituting \(\phi(\xi) = A e^{\mu \xi}\) into \Cref{eq:ODEx}, yields:
%
\begin{equation}\label{eq:ODEx2}
	\begin{split}
		\mu^4 + 2\alpha^2\left(\frac{D_{12}S_2}{D_{11}S_1}-2\frac{D_{66}S_3}{D_{11}S_1}\right)\mu^2 + \left(\alpha^4\frac{D_{22}S_4}{D_{11}S_1}-a^4\Omega_x^4\right) = 0.
	\end{split}
\end{equation}
%
The solution for \(\mu\) can be expressed as:
%
\begin{equation}\label{eq:mu}
	\begin{split}
		\mu_{1,2} = \pm\textit{i} \alpha_1, \qquad \mu_{3,4} = \pm \beta_1,
	\end{split}
\end{equation}
%
where,
%
\begin{subequations}\label{eq:alphax}
	\begin{align}
		\alpha_1 &= \alpha \sqrt{\sqrt{\left(\frac{D_{12}S_2}{D_{11}S_1}-2\frac{D_{66}S_3}{D_{11}S_1}\right)^2 - \frac{D_{22}S_4}{D_{11}S_1} + b^4\Omega_x^4} + \frac{D_{12}S_2}{D_{11}S_1} - 2\frac{D_{66}S_3}{D_{11}S_1}},\label{eq:alphax1}\\
		\beta_1 &= \alpha \sqrt{\sqrt{\left(\frac{D_{12}S_2}{D_{11}S_1}-2\frac{D_{66}S_3}{D_{11}S_1}\right)^2 - \frac{D_{22}S_4}{D_{11}S_1} + b^4\Omega_x^4} - \frac{D_{12}S_2}{D_{11}S_1} + 2\frac{D_{66}S_3}{D_{11}S_1}}.\label{eq:alphax2}
	\end{align}
\end{subequations}
%
The boundary conditions along the edges in the \(x\)-direction can be obtained from the remaining \(\delta \phi\) and \(\delta \frac{\id \phi}{\id \xi}\) parts in \Cref{eq:Rayleighx2}. 
The shear force equilibrium can be obtained from the \(\delta \phi\) part:
%
\begin{equation}\label{eq:inertial_forcex}
	\begin{split}
		&\left[ \left( \frac{4D_{66}}{a^2b^2} S_3 - \frac{D_{12}}{a^2b^2} S_2 \right) \frac{\id \phi}{\id \xi} 
		- \frac{D_{11}}{a^4} S_1 \frac{\id^3 \phi}{\id \xi^3} \right] \Big|^{\xi=1}_{\xi=-1}\\
		&\quad+ \frac{2k^v_{\xi=-1}}{a} S_1 \left(\phi\right)_{\xi=-1}
		 + \frac{2k^v_{\xi=1}}{a} S_1 \left(\phi\right)_{\xi=1} = 0,
	\end{split}
\end{equation}
%
and from the \(\delta \frac{\id \phi}{\id \xi}\) part, the bending moment equilibrium:
%
\begin{equation}\label{eq:inertial_forcex2}
	\begin{split}	
		\left( \frac{D_{12}}{a^2b^2} S_2 \phi + \frac{D_{11}}{a^4} S_1 \frac{\partial^2 \phi}{\partial \xi^2} \right) 
		\Big|^{\xi=1}_{\xi=-1} 
		&+ \frac{2k^r_{\xi=-1}}{a^3} S_1 \left(\frac{\partial \phi}{\partial \xi}\right)_{\xi=-1} \\
		&+ \frac{2k^r_{\xi=1}}{a^3} S_1 \left(\frac{\partial \phi}{\partial \xi}\right)_{\xi=1} = 0.
	\end{split}
\end{equation}
%
Thus, we can obtain the shear force and bending moment equilibrium along the edges \(\xi = -1\) and \(\xi = 1\) from \Cref{eq:inertial_forcex,eq:inertial_forcex2}, respectively, as:
%
\begin{subequations}\label{eq:boundaryx}
	\begin{align}
		&  \frac{\id^3 \phi}{\id \xi^3} - \alpha^2\left( \frac{4D_{66}S_3}{D_{11}S_1}  - \frac{D_{12}S_2}{D_{11}S_1}  \right) \frac{\id \phi}{\id \xi} 
		+\frac{2a^3k^v_{\xi=-1}}{D_{11}}\phi = 0, \quad &\xi = -1, \\
		& \frac{\id^2 \phi}{\id \xi^2} + \frac{\alpha^2 D_{12} S_2}{D_{11} S_1} \phi - \frac{2a k^r_{\xi=-1}}{D_{11}} \frac{\id \phi}{\id \xi} = 0, \quad &\xi = -1, \\
		& \frac{\id^3 \phi}{\id \xi^3} - \alpha^2\left( \frac{4D_{66}S_3}{D_{11}S_1}  - \frac{D_{12}S_2}{D_{11}S_1}  \right) \frac{\id \phi}{\id \xi} 
		-\frac{2a^3 k^v_{\xi=1}}{D_{11}} \phi = 0, \quad &\xi = 1, \\
		& \frac{\id^2 \phi}{\id \xi^2} + \frac{\alpha^2 D_{12} S_2}{D_{11} S_1} \phi + \frac{2a k^r_{\xi=1}}{D_{11}} \frac{\id \phi}{\id \xi} = 0, \quad &\xi = 1.
	\end{align}
\end{subequations}
%
Substituting \Cref{eq:phi_solution1} into \Cref{eq:boundaryx}, and denoting \( k^{v*}_{\xi} = \frac{2a^3 k^v_{\xi}}{D_{11}} \), \( k^{r*}_{\xi} = \frac{2a k^r_{\xi}}{D_{11}} \), \( \sin\alpha_1 = S_{\alpha_1} \), \( \cos\alpha_1 = C_{\alpha_1} \), \( \sinh\alpha_1 = Sh_{\alpha_1} \), \( \cosh\alpha_1 = Ch_{\alpha_1} \), \( \sin\beta_1 = S_{\beta_1} \), \( \cos\beta_1 = C_{\beta_1} \), \( \sinh\beta_1 = Sh_{\beta_1} \), and \( \cosh\beta_1 = Ch_{\beta_1} \), we have:
%
\begin{equation}\label{eq:ABx}
	\begin{split}
		&\left[ \begin{array}{c}
			\gamma_1 C_{\alpha_1} - k^{v*}_{\xi=-1} S_{\alpha_1} \\ 
			\gamma_3 S_{\alpha_1} + k^{r*}_{\xi=-1} \alpha_1 C_{\alpha_1} \\ 
			-\gamma_1 C_{\alpha_1} + k^{v*}_{\xi=1} S_{\alpha_1} \\ 
			\gamma_3 S_{\alpha_1} + k^{r*}_{\xi=1} \alpha_1 C_{\alpha_1}
		\end{array} 
		\begin{array}{c}
			\gamma_1 S_{\alpha_1} + k^{v*}_{\xi=-1} C_{\alpha_1} \\ 
			-\gamma_3 C_{\alpha_1} + k^{r*}_{\xi=-1} \alpha_1 S_{\alpha_1} \\ 
			\gamma_1 S_{\alpha_1} + k^{v*}_{\xi=1} C_{\alpha_1} \\ 
			\gamma_3 C_{\alpha_1} - k^{r*}_{\xi=1} \alpha_1 S_{\alpha_1}
		\end{array} 
		\begin{array}{c}
			\gamma_2 Ch_{\beta_1} - k^{v*}_{\xi=-1} Sh_{\beta_1} \\ 
			\gamma_4 Sh_{\beta_1} + k^{r*}_{\xi=-1} \beta_1 Ch_{\beta_1} \\ 
			-\gamma_2 Ch_{\beta_1} + k^{v*}_{\xi=1} Sh_{\beta_1} \\ 
			\gamma_4 Sh_{\beta_1} + k^{r*}_{\xi=1} \beta_1 Ch_{\beta_1}
		\end{array} \right. \\ 
		&\quad
		\left. \begin{array}{c}
			-\gamma_2 Sh_{\beta_1} + k^{v*}_{\xi=-1} Ch_{\beta_1} \\ 
			-\gamma_4 Ch_{\beta_1} - k^{r*}_{\xi=-1} \beta_1 Sh_{\beta_1} \\ 
			-\gamma_2 Sh_{\beta_1} + k^{v*}_{\xi=1} Ch_{\beta_1} \\ 
			\gamma_4 Ch_{\beta_1} + k^{r*}_{\xi=1} \beta_1 Sh_{\beta_1}
		\end{array} \right] \begin{Bmatrix}
			A_1 \\ A_2 \\ A_3 \\ A_4
		\end{Bmatrix} =  \begin{Bmatrix}
			0 \\ 0 \\ 0 \\ 0
		\end{Bmatrix},
	\end{split}
\end{equation}
%
or,
%
\begin{equation}\label{eq:ABx1}
		\mathbf{R}_x \mathbf{A} = \mathbf{0},
\end{equation}
%
where,
%
\begin{equation}\label{eq:gamma}
	\begin{split}
		&\gamma_1 = -\alpha_1^3 - \alpha^2 \left( \frac{4D_{66}S_3}{D_{11}S_1} - \frac{D_{12}S_2}{D_{11}S_1} \right) \alpha_1, \\
		&\gamma_2 = \beta_1^3 - \alpha^2 \left( \frac{4D_{66}S_3}{D_{11}S_1} - \frac{D_{12}S_2}{D_{11}S_1} \right) \beta_1, \\
		&\gamma_3 = -\alpha_1^2 + \frac{\alpha^2 D_{12} S_2}{D_{11} S_1}, \\
		&\gamma_4 = \beta_1^2 + \frac{\alpha^2 D_{12} S_2}{D_{11} S_1}.
	\end{split}
\end{equation}
%
Note that the classic boundary conditions can be obtained by selecting extremely large or small srping stiffness constants.
For non-trivial solutions, the characteristic equation or eigenvalue equation is obtained from the determinant of the matrix $\mathbf{R}_x$ in \Cref{eq:ABx1}, which must be zero.
However, solving these trancedental equations are cunmbersome and tedious, thus the DSM is introducted to avoid the ineffective computation.

To develop its dynamic stiffness matrix, with the help of \Cref{eq:phi_solution1}, the vertical displacement and rotation corresponding to the mode shape $\phi(\xi)$ along the $x$-direction at edges $\xi = -1$ and $\xi = 1$ can be expressed as:
%
\begin{equation}\label{eq:UBx}
	\begin{split}
		\begin{Bmatrix}
			\phi(\xi=-1)\\\frac{\id \phi(\xi=-1)}{\id \xi}\\\phi(\xi=1)\\\frac{\id \phi(\xi=1)}{\id \xi}
		\end{Bmatrix}
		= \begin{bmatrix}
			-S_{\alpha_1}&C_{\alpha_1}&-Sh_{\beta_1}&Ch_{\beta_1}\\
			\alpha_1 C_{\alpha_1}/a&\alpha_1S_{\alpha_1}/a&\beta_1Ch_{\beta_1}/a&-\beta_1Sh_{\beta_1}/a\\
			S_{\alpha_1}&C_{\alpha_1}&Sh_{\beta_1}&Ch_{\beta_1}\\
			\alpha_1 C_{\alpha_1}/a&-\alpha_1S_{\alpha_1}/a&\beta_1Ch_{\beta_1}/a&\beta_1Sh_{\beta_1}/a
		\end{bmatrix}
		\begin{Bmatrix}
			A_1\\A_2\\A_3\\A_4
		\end{Bmatrix} ,
	\end{split}
\end{equation}
%
or,
%
\begin{equation}\label{eq:UBx1}
	\mathbf{\delta}_x= \mathbf{Q}_x\mathbf{A}.
\end{equation}
%
Note that the eigenvector $\mathbf{A}$ can be expressed by multiplying the inverse matrix $\mathbf{Q}_x^{-1}$ on the left side of \Cref{eq:UBx1}, and then substituting $\mathbf{A}$ into \Cref{eq:ABx1}, we obtain:
%
\begin{equation}\label{eq:DSM_eqx}
	\mathbf{R}_x \mathbf{A}=\mathbf{R}_x\mathbf{Q}_x^{-1}\mathbf{\delta}_x = \mathbf{0}.
\end{equation}
%
where the dynamic stiffness matrix, denoted as $\mathbf{K}_x = \mathbf{R}_x\mathbf{Q}_x^{-1}$, can be obtained from \Cref{eq:DSM_eqx}. 
This matrix can be used to compute the natural frequencies of the system instead of solving the eigenvalue equation, 
and the method for the computation will be given in \Cref{sec:WWA}.

\subsection{Dynamic stiffness matrix corresponding to \(\omega_y\)}\label{sec:DSMy}
In this section, the mode shape \(\phi(\xi)\) derived in \(\Cref{sec:DSMx}\) is utilized to obtain the dynamic stiffness matrix in the \(y\)-direction.  
The vertical translational and rotational springs at \(\eta = -1\) are denoted as \(k^v_{\eta = -1}\) and \(k^r_{\eta = -1}\), respectively, while those at \(\eta = 1\) are represented by \(k^v_{\eta = 1}\) and \(k^r_{\eta = 1}\).

Similarly, the potential energy along the edges in the \( y \)-direction is given by:  
%
\begin{equation}\label{eq:poten_energy_edgey2}  
	\begin{split}  
		U^{III}_{mag} &= ab \int \left[ \frac{k^r_{\eta=-1}}{a^3} \left( \frac{\partial w}{\partial \eta} \right)^2 + \frac{k^v_{\eta=-1}}{a} w^2 \right]_{\eta=-1} \id \xi \\  
		&\quad + ab \int \left[ \frac{k^r_{\eta=1}}{a^3} \left( \frac{\partial w}{\partial \eta} \right)^2 + \frac{k^v_{\eta=1}}{a} w^2 \right]_{\eta=1} \id \xi.  
	\end{split}  
\end{equation} 
% 
The total potential energy of the plate in the \( y \)-direction can be determined from \Cref{eq:poten_energy_mag,eq:poten_energy_edgey2} as:
%
\begin{equation}\label{eq:poten_energy_toly}
	\begin{split}
		U_{mag}&=U^{I}_{mag}+U^{III}_{mag}\\
		&=\frac{ab}{2} \iint \left[ \frac{D_{11}}{a^4} \left( \frac{\partial^2 w}{\partial \xi^2} \right)^2 + \frac{2 D_{12}}{a^2 b^2} \frac{\partial^2 w}{\partial \xi^2} \frac{\partial^2 w}{\partial \eta^2} + \frac{D_{22}}{b^4} \left( \frac{\partial^2 w}{\partial \eta^2} \right)^2 \right. \\
		&\quad \left.  + \frac{4 D_{66}}{a^2 b^2}\left( \frac{\partial^2 w}{\partial \xi \partial \eta} \right)^2 \right] \, \id\xi \, \id\eta+ab \int \left[ \frac{k^r_{\eta=1}}{b^3} \left( \frac{\partial w}{\partial \eta} \right)^2 + \frac{k^v_{\eta=1}}{b} \left( w \right)^2 \right]_{\eta=1} \id\xi\\
		&\quad+ab \int \left[ \frac{k^r_{\eta=-1}}{b^3} \left( \frac{\partial w}{\partial \eta} \right)^2 + \frac{k^v_{\eta=-1}}{b} \left( w \right)^2 \right]_{\eta=-1} \id\xi.
	\end{split}
\end{equation}

By substituting \Cref{eq:displace_func} into \Cref{eq:poten_energy_toly}, we obtain:  
%
\begin{equation}\label{eq:poten_energy_toly2}
	\begin{split}
		U_{mag} &= U^{I}_{mag} + U^{III}_{mag} \\
		&= \frac{ab}{2} \int_{-1}^{1} \left[ \frac{D_{11}}{a^4} T_4 \psi^2 
		+ \frac{2D_{12}}{a^2b^2} T_2 \frac{\id^2\psi }{\id \eta^2} \psi 
		+ \frac{D_{22}}{b^4} T_1 \left(\frac{\id^2\psi }{\id \eta^2} \right)^2 \right. \\
		&\quad \left. + \frac{4D_{66}}{a^2b^2} T_3 \left(\frac{\id \psi }{\id \eta} \right)^2 \right] \, \id\eta  + ab \left[ \frac{k^r_{\eta=1}}{b^3} T \left( \frac{\id \psi}{\id \eta} \right)^2
		+ \frac{k^v_{\eta=1}}{b} T_1 \left(\psi\right)^2 \right]_{\eta=1} \\
		&\quad + ab \left[ \frac{k^r_{\eta=-1}}{b^3} T \left( \frac{\id \psi}{\id \eta} \right)^2
		+ \frac{k^v_{\eta=-1}}{b} T_1 \left(\psi\right)^2 \right]_{\eta=-1},
	\end{split}
\end{equation}
%
where the integral parameters are defined as:  
%
\begin{equation}\label{eq:inte_psi}  
	\begin{split}  
		&T_1 = \int_{-1}^{1} \phi^2 \, \id\xi, \\  
		&T_2 = \int_{-1}^{1} \left( \frac{\id^2\phi}{\id\xi^2} \phi \right) \, \id\xi, \\  
		&T_3 = \int_{-1}^{1} \left( \frac{\id\phi}{\id\xi} \right)^2 \, \id\xi, \\  
		&T_4 = \int_{-1}^{1} \left( \frac{\id^2\phi}{\id\xi^2} \right)^2 \, \id\xi.  
	\end{split}  
\end{equation}  
%
The coefficient \( T_0 \) of the kinetic energy in \Cref{eq:kine_energy_mag} for the plate can be expressed as:
%
\begin{equation}\label{eq:kine_energyy}
	\begin{split}
		T_0 &= \frac{ab}{2} \rho h T_1 \int_{-1}^{1} \psi^2 \, \id\eta.
	\end{split}
\end{equation}

Taking the Rayleigh principle in the form:
%
\begin{equation}\label{eq:Rayleighy}
	\begin{split}
		\delta U_{mag} &= \omega_y^2 \, \delta T_0.
	\end{split}
\end{equation}
%
By substituting \Cref{eq:poten_energy_toly2,eq:kine_energyy} into \Cref{eq:Rayleighy}, and applying the variational calculus to eliminate \( \delta \psi \) and \( \delta\frac{\id \psi}{\id \eta} \) in \Cref{eq:Rayleighy}, we obtain:
%
\begin{equation}\label{eq:Rayleighy2}
	\begin{split}
		0 &= \int_{-1}^{1} \left[ \frac{D_{22}}{b^4} T_1 \frac{d^4 \psi}{\id \eta^4} 
		+ \left( \frac{2D_{12}}{a^2b^2} T_2 - \frac{4D_{66}}{a^2b^2} T_3 \right) \frac{\id^2 \psi}{\id \eta^2} \right. \\
		&\quad \left. + \left( \frac{D_{11}}{a^4} T_4 - \omega_y^2 \rho h T_1 \right) \psi \right] \delta \psi \, \id\eta \\
		&\quad + \frac{2k^v_{\eta=-1}}{b} T_1 (\psi \delta \psi)_{\eta=-1} 
		+ \frac{2k^v_{\eta=1}}{b} T_1 (\psi \delta \psi)_{\eta=1} \\
		&\quad + \left[ \left( \frac{4D_{66}}{a^2b^2} T_3 - \frac{D_{12}}{a^2b^2} T_2 \right) \frac{\id \psi}{\id \eta} 
		- \frac{D_{22}}{b^4} T_1 \frac{\id^3 \psi}{\id \eta^3} \right] \delta \psi \Big|^{\eta=1}_{\eta=-1} \\
		&\quad + \left( \frac{D_{12}}{a^2b^2} T_2 \psi + \frac{D_{22}}{b^4} T_1 \frac{\id^2 \psi}{\id \eta^2} \right) 
		\delta\frac{\id  \psi}{\id \eta} \Big|^{\eta=1}_{\eta=-1} \\
		&\quad + \frac{2k^r_{\eta=-1}}{b^3} T_1 \left( \frac{\id \psi}{\id \eta} \delta\frac{\id \psi}{\id \eta} \right)_{\eta=-1} 
		+ \frac{2k^r_{\eta=1}}{b^3} T_1 \left( \frac{\id \psi}{\id \eta} \delta\frac{\id \psi}{\id \eta} \right)_{\eta=1}.
	\end{split}
\end{equation}
%
Thus, the governing differential equation in the \( y \)-direction can be obtained from the integration part in \Cref{eq:Rayleighy2}:
%
\begin{equation}\label{eq:ODEy}
		\frac{\id^4 \psi}{\id \eta^4} + \frac{2}{\alpha^2}\left(\frac{D_{12}T_2}{D_{22}T_1} - 2\frac{D_{66}T_3}{D_{22}T_1}\right)\frac{\id^2 \psi}{\id \eta^2} + \left( \frac{D_{11}T_4}{\alpha^4D_{22}T_1} - \frac{b^4D_{11}}{D_{22}} \Omega_y^4 \right) \psi = 0,
\end{equation}
%
where \( \Omega_y = \sqrt[4]{\frac{\omega_y^2 \rho h}{D_{11}}} \).
Substituting \( \psi(\eta) = B e^{\lambda \eta} \) into \Cref{eq:ODEy}, yields:
%
\begin{equation}\label{eq:ODEy2}
	\begin{split}
		\lambda^4 + \frac{2}{\alpha^2} \left( \frac{D_{12}T_2}{D_{22}T_1} - 2\frac{D_{66}T_3}{D_{22}T_1} \right) \lambda^2 + \left( \frac{D_{11}T_4}{\alpha^4D_{22}T_1} - \frac{b^4D_{11}}{D_{22}} \Omega_y^4 \right) = 0.
	\end{split}
\end{equation}
%
The solution for \( \lambda \) can be expressed as:
%
\begin{equation}\label{eq:lambda}
	\begin{split}
		\lambda_{1,2} &= \pm \textit{i} \alpha_2, \qquad \lambda_{3,4} = \pm \beta_2,
	\end{split}
\end{equation}
%
where,
%
\begin{subequations}\label{eq:beta}
	\begin{align}
		\alpha_2 &= \frac{1}{\alpha} \sqrt{ \sqrt{ \left( \frac{D_{12}T_2}{D_{22}T_1} - 2\frac{D_{66}T_3}{D_{22}T_1} \right)^2 - \frac{D_{11}T_4}{D_{22}T_1} + \frac{a^4 D_{11}}{D_{22}} \Omega_y^4 } + \frac{D_{12}T_2}{D_{22}T_1} - 2 \frac{D_{66}T_3}{D_{22}T_1}}, \\
		\beta_2 &= \frac{1}{\alpha} \sqrt{ \sqrt{ \left( \frac{D_{12}T_2}{D_{22}T_1} - 2 \frac{D_{66}T_3}{D_{22}T_1} \right)^2 - \frac{D_{11}T_4}{D_{22}T_1} + \frac{a^4 D_{11}}{D_{22}} \Omega_y^4 } - \frac{D_{12}T_2}{D_{22}T_1} + 2 \frac{D_{66}T_3}{D_{22}T_1}}.
	\end{align}
\end{subequations}
%
The boundary conditions along the edges in the \( y \)-direction can be obtained from the remaining \( \delta \psi \) and \( \delta \frac{\id \psi}{\id \eta} \) parts in \Cref{eq:Rayleighy2}. 
The shear force equilibrium can be obtained from the \( \delta \psi \) part:
%
\begin{equation}\label{eq:inertial_forcey}
	\begin{split}
		& \left[ \left( \frac{4D_{66}}{a^2b^2} T_3 - \frac{D_{12}}{a^2b^2} T_2 \right) \frac{\id \psi}{\id \eta} 
		- \frac{D_{22}}{b^4} T_1 \frac{\id^3 \psi}{\id \eta^3} \right] \Big|^{\eta=1}_{\eta=-1} \\
		&\quad + \frac{2k^v_{\eta=-1}}{b} T_1 (\psi )_{\eta=-1} + \frac{2k^v_{\eta=1}}{b} T_1 (\psi )_{\eta=1} = 0,
	\end{split}
\end{equation}
%
and from the \( \delta \frac{\id \psi}{\id \eta} \) part, the bending moment equilibrium:
%
\begin{equation}\label{eq:inertial_forcey2}	
 \begin{split}
 	\left( \frac{D_{12}}{a^2b^2} T_2 \psi + \frac{D_{22}}{b^4} T_1 \frac{\id^2 \psi}{\id \eta^2} \right) 
 	\Big|^{\eta=1}_{\eta=-1} 
 	&+ \frac{2k^r_{\eta=-1}}{b^3} T_1 \left( \frac{\id \psi}{\id \eta} \right)_{\eta=-1} \\
 	&+ \frac{2k^r_{\eta=1}}{b^3} T_1 \left( \frac{\id \psi}{\id \eta} \right)_{\eta=1}=0.
 \end{split}
\end{equation}
%
Thus, we can obtain the shear force and bending moment equilibrium along the edges \( \eta = -1 \) and \( \eta = 1 \) from \Cref{eq:inertial_forcey,eq:inertial_forcey2}, respectively, as:
%
\begin{subequations}\label{eq:boundaryy}
	\begin{align}
		&  \frac{\id^3 \psi}{\id \eta^3}-\left( \frac{4D_{66}T_3}{\alpha^2D_{22}T_1}  - \frac{D_{12}T_2}{\alpha^2D_{22}T_1}  \right) \frac{\id \psi}{\id \eta} 
		+\frac{2b^3k^v_{\eta=-1}}{D_{22}}\psi=0,\quad &\eta=-1,\\
		& \frac{\id^2 \psi}{\id \eta^2}+\frac{D_{12}T_2}{\alpha^2D_{22}T_1} \psi -\frac{2bk^r_{\eta=-1}}{D_{22}}\frac{\id\psi }{\id \eta}=0,\quad &\eta=-1,\\
		&  \frac{\id^3 \psi}{\id \eta^3}-\left( \frac{4D_{66}T_3}{\alpha^2D_{22}T_1}  - \frac{D_{12}T_2}{\alpha^2D_{22}T_1}  \right) \frac{\id \psi}{\id \eta} 
		-\frac{2b^3k^v_{\eta=1}}{D_{22}}\psi=0,\quad &\eta=1,\\
		&\frac{\id^2 \psi}{\id \eta^2}+\frac{D_{12}T_2}{\alpha^2D_{22}T_1} \psi +\frac{2bk^r_{\eta=1}}{D_{22}}\frac{\id\psi }{\id \eta}=0,\quad &\eta=1.
	\end{align}
\end{subequations}
%
Substituting \Cref{eq:phi_solution2} into \Cref{eq:boundaryy} and introducing the notations 
\( k^{v*}_{\eta} = \frac{2b^3 k^v_{\eta}}{D_{22}} \), 
\( k^{r*}_{\eta} = \frac{2b k^r_{\eta}}{D_{22}} \), 
\( \sin\alpha_2 = S_{\alpha_2} \), 
\( \cos\alpha_2 = C_{\alpha_2} \), 
\( \sinh\alpha_2 = Sh_{\alpha_2} \), 
\( \cosh\alpha_2 = Ch_{\alpha_2} \), 
\( \sin\beta_2 = S_{\beta_2} \), 
\( \cos\beta_2 = C_{\beta_2} \), 
\( \sinh\beta_2 = Sh_{\beta_2} \), and 
\( \cosh\beta_2 = Ch_{\beta_2} \), 
We obtain:
%
\begin{equation}\label{eq:ABy}
	\begin{split}
		&\left[ \begin{array}{c}
			\hat{\gamma}_1C_{\alpha_2}-k^{v*}_{\eta=-1}S_{\alpha_2} \\ \hat{\gamma}_3S_{\alpha_2}+k^{r*}_{\eta=-1}\alpha_2C_{\alpha_2} \\ -\hat{\gamma}_1C_{\alpha_2}+k^{v*}_{\eta=1}S_{\alpha_2} \\ \hat{\gamma}_3S_{\alpha_2}+k^{r*}_{\eta=1}\alpha_2C_{\alpha_2}
		\end{array} \begin{array}{c}
			\hat{\gamma}_1S_{\alpha_2}+k^{v*}_{\eta=-1}C_{\alpha_2} \\ -\hat{\gamma}_3C_{\alpha_2}+k^{r*}_{\eta=-1}\alpha_2S_{\alpha_2} \\ \hat{\gamma}_1S_{\alpha_2}+k^{v*}_{\eta=1}C_{\alpha_2} \\ \hat{\gamma}_3C_{\alpha_2}-k^{r*}_{\eta=1}\alpha_2S_{\alpha_2}
		\end{array} \begin{array}{c}
			\hat{\gamma}_2Ch_{\beta_2}-k^{v*}_{\eta=-1}Sh_{\beta_2} \\ \hat{\gamma}_4Sh_{\beta_2}+k^{r*}_{\eta=-1}\beta_2Ch_{\beta_2} \\ -\hat{\gamma}_2Ch_{\beta_2}+k^{v*}_{\eta=1}Sh_{\beta_2} \\ \hat{\gamma}_4Sh_{\beta_2}+k^{r*}_{\eta=1}\beta_2Ch_{\beta_2}
		\end{array} \right. \\
		&\quad
		\left. \begin{array}{c}
			-\hat{\gamma}_2Sh_{\beta_2}+k^{v*}_{\eta=-1}Ch_{\beta_2} \\ -\hat{\gamma}_4Ch_{\beta_2}-k^{r*}_{\eta=-1}\beta_2Sh_{\beta_2} \\ -\hat{\gamma}_2Sh_{\beta_2}+k^{v*}_{\eta=1}Ch_{\beta_2} \\ \hat{\gamma}_4Ch_{\beta_2}+k^{r*}_{\eta=1}\beta_2Sh_{\beta_2}
		\end{array} \right]\begin{Bmatrix}
			B_1\\B_2\\B_3\\B_4
		\end{Bmatrix}=\begin{Bmatrix}
			0\\0\\0\\0
		\end{Bmatrix},
	\end{split}
\end{equation}
%
or,
%
\begin{equation}\label{eq:ABy1}
	\mathbf{R}_y \mathbf{B} = \mathbf{0},
\end{equation}
%
where,
%
\begin{equation}\label{eq:gammay}
	\begin{split}
		&\hat{\gamma}_1=-\alpha_2^3-\left( \frac{4D_{66}T_3}{\alpha^2D_{22}T_1}  - \frac{D_{12}T_2}{\alpha^2D_{22}T_1}  \right)\alpha_2,\\
		&\hat{\gamma}_2=\beta_2^3-\left( \frac{4D_{66}T_3}{\alpha^2D_{22}T_1}  - \frac{D_{12}T_2}{\alpha^2D_{22}T_1}  \right)\beta_2,\\
		&\hat{\gamma}_3=-\alpha_2^2+\frac{D_{12}T_2}{\alpha^2D_{22}T_1},\\
		&\hat{\gamma}_4=\beta_2^2+\frac{D_{12}T_2}{\alpha^2D_{22}T_1}.
	\end{split}
\end{equation}
%
With the help of \Cref{eq:phi_solution2}, the vertical displacement and rotation corresponding to the mode shape $\psi$ along the $y$-direction at the edges $\eta = -1$ and $\eta = 1$ can be expressed as:
%
\begin{equation}\label{eq:UBy}
	\begin{split}
		\begin{Bmatrix}
			\psi_{\eta=-1}\\\frac{\id \psi}{\id \eta}_{\eta=-1}\\\psi_{\eta=1}\\\frac{\id \psi}{\id \eta}_{\eta=1}
		\end{Bmatrix}
		= \begin{bmatrix}
			-S_{\alpha_2} & C_{\alpha_2} & -Sh_{\beta_2} & Ch_{\beta_2} \\
			\frac{\alpha_2 C_{\alpha_2}}{b} & \frac{\alpha_2 S_{\alpha_2}}{b} & \frac{\beta_2 Ch_{\beta_2}}{b} & -\frac{\beta_2 Sh_{\beta_2}}{b} \\
			S_{\alpha_2} & C_{\alpha_2} & Sh_{\beta_2} & Ch_{\beta_2} \\
			\frac{\alpha_2 C_{\alpha_2}}{b} & -\frac{\alpha_2 S_{\alpha_2}}{b} & \frac{\beta_2 Ch_{\beta_2}}{b} & \frac{\beta_2 Sh_{\beta_2}}{b}
		\end{bmatrix}
		\begin{Bmatrix}
			B_1\\B_2\\B_3\\B_4
		\end{Bmatrix},		
	\end{split}
\end{equation}
%
or,
%
\begin{equation}\label{eq:UBy1}
	\mathbf{\delta}_y= \mathbf{Q}_y\mathbf{B}.
\end{equation}
%
Note that the eigenvector \(\mathbf{B}\) can be expressed by multiplying the inverse matrix \(\mathbf{Q}_y^{-1}\) on the left-hand side of \Cref{eq:UBy1}, and then substituting \(\mathbf{B}\) into \Cref{eq:ABy1}, we obtain:
%
\begin{equation}\label{eq:DSM_eqy}
	\mathbf{R}_y\mathbf{B}=\mathbf{R}_y \mathbf{Q}_x^{-1} 
	\mathbf{\delta}_y = \mathbf{0},
\end{equation}
%
where the dynamic stiffness matrix, denoted as \(\mathbf{K}_y = \mathbf{R}_y \mathbf{Q}_y^{-1}\), can be obtained from \Cref{eq:DSM_eqy}.

\section{Frequency and mode shape computation}\label{sec:WWA}
\subsection{Wittrick–Williams algorithm}\label{sec:WWA2}
The Wittrick–Williams (W-W) algorithm \cite{wittrick1971general} is an effective method for determining the natural frequencies from the dynamic stiffness matrix with high reliability. 
Instead of directly solving the equations, the algorithm computes the total number \(J\) of natural frequencies below a given frequency \(\omega^*\), which is represented as:
\begin{equation}\label{eq:WWalgorithm}
	J(\omega^*) = J_0(\omega^*) + s\{\mathbf{K}^{\Delta}(\omega^*)\} = J_0(\omega^*) + J_k(\omega^*),
\end{equation}
where \(J_0\) represents the number of natural frequencies for the system with both ends fully clamped, \(\mathbf{K}^{\Delta}\) is the upper triangular matrix obtained from the dynamic stiffness matrix \(\mathbf{K}\) after applying Gaussian elimination, and \(J_k(\omega^*)\) denotes the number of negative elements in the leading diagonal of \(\mathbf{K}^{\Delta}\).

It should be noted that the \( J_0 \) count is a crucial aspect when applying the W-W algorithm. Many previous studies use a sufficiently fine mesh or enough terms in series expansions to capture all fully clamped natural frequencies, ensuring computational accuracy \cite{banerjee2015dynamic}. However, this approach can make the application process cumbersome. 
To address this issue, the fully clamped problem can be replaced with a simply supported problem, where the Navier solution for the simply supported plate is used to count \( J_0 \) \citep{liu2015exact}. 
Nevertheless, since analytical solutions in DSM methods involve an infinite series of Fourier terms, a sufficient number of truncation terms is required to ensure accuracy and convergence.

However, the idea of solving the simply supported problem proposes an effective and systematic approach to indirectly determine the \( J_0 \) count for a fully clamped structure using the 'hypothetical structure method', where the boundary conditions are modeled as pinned supports rather than clamped ones \cite{han2018improved}:
%
\begin{equation}\label{eq:J_o_pin}
	J_0(p_1,\omega^*) = J(\bar{p}_1,\omega^*) - J_k(\bar{p}_1,\omega^*),
\end{equation}
%
where \( p_1 \) and \( \bar{p}_1 \) represent clamped and pinned supports, respectively.
By substituting \Cref{eq:J_o_pin} into \Cref{eq:WWalgorithm} we get the algorithm as:  
%
\begin{equation}\label{eq:new_WWalgorithm}
	J(p,\omega^*) = J(\bar{p}_1,\omega^*) - J_k(\bar{p}_1,\omega^*) + J_k(p,\omega^*)
\end{equation} 
%
where \( p \) represents the original boundary conditions of the structure.
Therefore, the challenge of determining \( J_0(p_1, \omega^*) \) can be transformed into the problem of solving \( J(\bar{p}_1, \omega^*) \) instead. The eigenvalue equation corresponding to the natural frequency parameter \( \Omega_x \) can be obtained from the determinant of the coefficient matrix \( \mathbf{R}_x \) in \Cref{eq:ABx}, which is given by:
%
\begin{equation}\label{eq:SSx} 
	\sin 2\alpha_1 = 0.
\end{equation}
%
With the help of \Cref{eq:alphax1,eq:SSx}, the closed-form solution of the \(n_x\)th simply supported frequency \(\Omega_{x,n_x}\) for the given \(n_y\)-order \(\psi_{n_y}(\eta)\) can be expressed as:
%
\begin{equation}\label{eq:J_simpx}
	\Omega_{x,n_x} =
	\frac{1}{b} \sqrt[4]{%
		%\large % slightly smaller than \Large
		\begin{aligned}
			&\left[\left(\frac{n_x\pi}{2\alpha}\right)^2 - \frac{D_{12}S_2}{D_{11}S_1} + 2 \frac{D_{66}S_3}{D_{11}S_1}\right]^2 - \left(\frac{D_{12}S_2}{D_{11}S_1} - 2 \frac{D_{66}S_3}{D_{11}S_1}\right)^2 \\
			&+ \frac{D_{22}S_4}{D_{11}S_1}.
		\end{aligned}
	}
\end{equation}
%
For \(\Omega_{x,n_x} \leq \Omega_x^* < \Omega_{x,n_{x+1}}\), \(J(\bar{p}_1, \Omega_x^*) = n_x\).
Similarly, the closed-form solution of the \(n_y\)th simply supported frequency \(\Omega_{y,n_y}\) for the given \(n_x\)-order \(\phi_{n_x}(\xi)\) can be expressed as:
%
\begin{equation}\label{eq:J_simpy}
	\Omega_{y,n_y}= 
	\frac{1}{a} \sqrt[4]{%
		\begin{aligned}
			&\frac{D_{22}}{D_{11}}\left\{\left[\left(\frac{n_y\pi\alpha}{2}\right)^2 - \frac{D_{12}T_2}{D_{22}T_1} + 2 \frac{D_{66}T_3}{D_{22}T_1}\right]^2 - \left(\frac{D_{12}T_2}{D_{22}T_1} - 2 \frac{D_{66}T_3}{D_{22}T_1}\right)^2\right. \\
			&\left.  + \frac{D_{11}T_4}{D_{22}T_1}\right\}.
		\end{aligned}
	}
\end{equation}
%
For \(\Omega_{y,n_y} \leq \Omega_y^* < \Omega_{y,n_{y+1}}\), \(J(\bar{p}_1, \Omega_y^*) = n_y\).
According to the relationships 
$\Omega_x = \sqrt[4]{\frac{\omega_x^2 \rho h}{D_{11}}}$ and\ $\Omega_y = \sqrt[4]{\frac{\omega_y^2 \rho h}{D_{11}}}$, 
the values of \(J(\bar{p}_1, \omega_x^*)\) and \(J(\bar{p}_1, \omega_y^*)\) can be derived from \(J(\bar{p}_1, \Omega_x^*)\) and \(J(\bar{p}_1, \Omega_y^*)\), respectively. Therefore, this refined W-W algorithm can be applied to estimate the lower and upper bounds of the frequency range, denoted as \(\omega_l\) and \(\omega_u\), yielding an approximation for the frequency \(\omega_a \in (\omega_l, \omega_u)\).

The mode shape coefficients \( A_1 \) to \( A_4 \) and \( B_1 \) to \( B_4 \) in the eigenvectors \( \mathbf{A} \) and \( \mathbf{B} \) for all classic boundary conditions are provided in \cite{xing2020extended}. 
Alternatively, these coefficients can also be obtained through a simple numerical method, which this work presents as an approach.
Here, we illustrate solving the eigenvector \( \mathbf{A} \) as an example.
By assuming the exact natural frequency as \( \omega_k \), we can expand the coefficient matrix \( \mathbf{R}_{x} \) in \Cref{eq:ABx} using a first-order Taylor series about \( \omega_a \):
%
\begin{equation}\label{eq:Taylor}
	\mathbf{R}_{x,k}(\omega_k)\mathbf{A}_k = \mathbf{R}_{x,a}\mathbf{A}_k + (\omega_k - \omega_a)\mathbf{R}'_{x,a}\mathbf{A}_k + O\left((\omega_k - \omega_a)^2\right) = 0.
\end{equation}
%
Ignoring higher-order terms, an eigenvalue problem can be derived from \Cref{eq:Taylor}:
%
\begin{equation}\label{eq:Newgeigen}
	(\mathbf{R}^{'}_{x,a})^{-1} \mathbf{R}_{x,a} \mathbf{A} = (\omega_a - \omega_k)\mathbf{A} = \tau \mathbf{A}.
\end{equation}
%
This eigenvalue problem can be solved using the inverse iteration procedure \cite{yuan2004second}:
%
\begin{equation}\label{eq:Inversepower}
	\bar{\mathbf{A}}^{(i+1)} = \mathbf{R}_{x,a}^{-1} \mathbf{R}^{'}_{x,a} \mathbf{A}^{(i)},	
\end{equation}
%
where the initial guess for \( \mathbf{A}^{(0)} \) is a column vector consisting of four randomly generated elements, each of which falls within the range (0,1). The updated eigenvalue for the next step can be obtained as:
%
\begin{equation}\label{eq:Newgeigenvalue}
	\tau^{(i+1)} = \frac{1}{\bar{A}_j^{(i+1)}},
\end{equation}
%
where,
%
\begin{equation}\label{eq:updatetau}
	\mid \bar{A}_j^{(i+1)} \mid = \max(\mid \bar{A}_1^{(i+1)} \mid, \mid \bar{A}_2^{(i+1)} \mid, \mid\bar{A}_3^{(i+1)}\mid , \mid \bar{A}_4^{(i+1)} \mid).
\end{equation}
%
The updated eigenvector can be obtained as:
%
\begin{equation}\label{eq:updateeigenvalue}
	\mathbf{A}^{(i+1)} = \tau^{(i+1)} \bar{\mathbf{A}}^{(i+1)}.
\end{equation}
%
The procedure can be controlled by the error tolerance \( \epsilon \) or maximum allowed steps \( i_{\text{max}} \):
%
\begin{subequations}
	\begin{align}
		\max \mid A_n^{(i+1)} - A_n^{(i)} \mid < \epsilon, \\ 
		i = i_{\text{max}}.
	\end{align}
\end{subequations}
%
Note that the mode shape coefficients \( A_1 \) to \( A_4 \) obtained from \(\mathbf{A}^{(i+1)}\) are applied for the elastically restrained boundary conditions.

\subsection{Application procedure}
The procedure of the proposed method is as follows:   
\subsubsection{Iterative procedure}\label{sec:Iterative_procedure}
\begin{itemize}[]
	\item \textbf{Step 1} Assume initial integral parameters \(S^{(0)}_1, S^{(0)}_2, S^{(0)}_3\), and \(S^{(0)}_4\) in the \(y\)-direction. Using the given boundary conditions (BCs) at \(\xi = -1\) and \(\xi = 1\), determine \(\mathbf{K}^{(0)}_x\) from \Cref{eq:DSM_eqx}. Then, apply the computational algorithms in \Cref{sec:WWA2} to compute the \(n_x\)-th frequency \(\omega^{(0)}_{x,n_x}\) and its corresponding mode shape \(\phi^{(0)}_{n_x}\), where \(n_x = 1, 2, 3, \ldots\).

	\item \textbf{Step 2} Use \(\phi^{(0)}_{n_x}\) as the prescribed mode to determine \(\mathbf{K}^{(1)}_y\) in \Cref{eq:DSM_eqy}, considering the BCs at \(\eta = -1\) and \(\eta = 1\). Apply the computational algorithms to obtain the \(n_y\)-th frequency \(\omega^{(1)}_{y,n_y}\) and its corresponding mode shape \(\psi^{(1)}_{n_y}\), where \(n_y = 1, 2, 3, \ldots\). This completes the first iteration cycle.

	\item \textbf{Step 3} Use \(\psi^{(1)}_{n_y}\) as the prescribed \(n_y\)-th mode shape in the \(y\)-direction to compute \(\mathbf{K}^{(1)}_x\) from \Cref{eq:DSM_eqx}, then determine the \(n_x\)-th frequency \(\omega^{(1)}_{x,n_x}\) and its corresponding mode shape \(\phi^{(1)}_{n_x}\).

	\item \textbf{Step 4} Use \(\phi^{(1)}_{n_x}\) as the prescribed mode in the \(x\)-direction to compute the \(n_y\)-th frequency \(\omega^{(2)}_{y,n_y}\) and its corresponding mode shape \(\psi^{(2)}_{n_y}\), completing the second iteration cycle.

	\item \textbf{Step 5} Stop the iteration if \(\mid\omega^{(i)}_{x,n_x} - \omega^{(i+1)}_{x,n_x}\mid \leq \Delta\omega\) or \(\mid\omega^{(i)}_{y,n_y} - \omega^{(i+1)}_{y,n_y}\mid \leq \Delta\omega\), where \(\Delta\omega\) is the frequency interval defined in \Cref{sec:WWA}, where, \(\Delta\omega = \omega_u - \omega_l\).

	\item \textbf{Step 6} Finally, construct the \((n_x, n_y)\)-th mode shape as \(w(\xi, \eta) = \phi_{n_x}(\xi) \psi_{n_y}(\eta)\) using \Cref{eq:displace_func}.  
\end{itemize}
\subsubsection{Non-iterative Procedure}\label{sec:Noniterative_procedure}  
Although the iterative procedure in \Cref{sec:Iterative_procedure} can be used to obtain any desired mode, a non-iterative approach can enhance the efficiency of numerical calculations when an exact mode is known.

Once a nonzero mode shape \(\phi_{n_x}(\xi) \psi_{n_y}(\eta)\) is determined by following the iterative procedure in \Cref{sec:Iterative_procedure}, \(\psi_{n_y}(\eta)\) can be chosen as the exact known mode in the \(y\)-direction.  
This allows the computation of infinitely many modes \(\phi(\xi)\) in the \(x\)-direction by following only \textbf{Step 1}, without requiring further iterations.  
Similarly, if \(\phi_{n_x}(\xi)\) is selected as the exact known mode in the \(x\)-direction, infinitely many modes \(\psi(\eta)\) in the \(y\)-direction can be computed by following only \textbf{Step 2}, without iterations.

For example, if the \((n_x, n_y)\)-th mode is nonzero and obtained through the iteration procedure, then based on \(\psi_{n_y}(\eta)\), the modes \((1,n_y)\), \((2,n_y)\), ..., \((\infty,n_y)\) of the plate can be computed without further iteration.  
Similarly, based on \(\phi_{n_x}(\xi)\), the modes \((n_x,1)\), \((n_x,2)\), ..., \((n_x,\infty)\) of the plate can be determined.  
That is, only one iteration procedure is required to calculate all frequencies of the plate, which helps reduce the computational cost.

\FloatBarrier
\section{Numerical Results}  
This section presents the numerical validation of the proposed method for classic boundary conditions and elastically restrained boundary conditions.  
For all numerical calculations, the initial integral parameters are assumed as \(S^{(0)}_1=1\), \(S^{(0)}_2=1\), \(S^{(0)}_3=1\), and \(S^{(0)}_4=10\) in the \(y\)-direction, serving as the starting point of \textbf{Step 1} for any mode in all BCs.  
In this section, the interval between the upper and lower bounds of the non-dimensional frequency parameter, \(2a\Delta\Omega\), is set to 0.005, defining the error range.  
According to our numerical calculations, two iteration cycles are sufficient to meet the convergence requirement for most cases, with at most three cycles needed when applying the iterative procedure in \Cref{sec:Iterative_procedure}.

\subsection{Classical boundary conditions}
In this subsection, the proposed method is validated by comparison with the extended SOV method \cite{xing2020extended}.  
The properties of the orthotropic plate, consistent with those in \cite{xing2020extended}, are as follows:  
\(E_1 = 185 \nugpa\), \(E_2 = 10.5 \nugpa\), \(G_{12} = 7.3 \nugpa\), \(\rho = 1600 \urho\), and \(\nu_{12} = 0.28\).

The initial guess for \(S^{(0)}_1, S^{(0)}_2, S^{(0)}_3\), and \(S^{(0)}_4\) is assumed to be \(1, 1, 1, 10\) for all boundary conditions and mode orders. This differs from the iterative SOV method \citealp{xing2018overall}, which requires a different initial guess for each mode order.  
The translational springs (\(k^v_\xi\)) and rotational springs (\(k^r_\xi\)) along all edges can be set to zero or infinity (represented as \(1 \times 10^{15} \ustif\) in the numerical calculations of this study) to obtain different classic boundary conditions.

The results for SSSS, SCSF, GCGC, CCCC, SSCC, SCCC, GGCC, CCFF, CFCF, CFFF, and FFFF boundary conditions are presented in \Cref{tab:sov1,tab:sov2,tab:sov3}. These results demonstrate high accuracy compared to the extended SOV method, with errors remaining smaller than the interval between the upper and lower bounds (0.005). The frequency parameters in both directions are equal (\(2a\Omega_x - 2a\Omega_y = 0\)) in almost all cases, with a few exceptions where \(2a\Omega_x - 2a\Omega_y = 0.005\).  
\Cref{fig:FFFF} shows the first six nonzero mode shapes of a square orthotropic plate with FFFF boundary conditions.

\begin{figure}
	\centering
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\textwidth]{figs/FFFF1.png}
		\caption[]%
		{{\small }}    
		\label{fig:FFF1}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.49\textwidth}  
		\centering 
		\includegraphics[width=\textwidth]{figs/FFFF2.png}
		\caption[]%
		{{\small }}    
		\label{fig:FFF2}
	\end{subfigure}
	\vfill
	\begin{subfigure}[b]{0.49\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{figs/FFFF3.png}
		\caption[]%
		{{\small }}    
		\label{fig:FFF3}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.49\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{figs/FFFF4.png}
		\caption[]%
		{{\small }}    
		\label{fig:FFF4}
	\end{subfigure}
	\vskip\baselineskip
	\begin{subfigure}[b]{0.49\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{figs/FFFF5.png}
		\caption[]%
		{{\small}}    
		\label{fig:FFF5}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.49\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{figs/FFFF6.png}
		\caption[]%
		{{\small }}    
		\label{fig:FFF6}
	\end{subfigure}
	\caption[]  
	{\small The first six nonzero mode shapes of a square orthotropic plate with FFFF boundary conditions: (a) the first mode; (b) the second mode; (c) the third mode; (d) the fourth mode; (e) the fifth mode; (f) the sixth mode.}  
	\label{fig:FFFF}
\end{figure}

\begin{table}[!htbp] 
	\centering
	\caption{The first seven frequency parameter $2a\Omega$ of of orthotropic rectangular plates with SSSS, SCSF and GCGC boundary conditions.}
	\begin{tabular}{c c l c c c c c c c}
		\toprule
		\multicolumn{3}{c}{} & \multicolumn{7}{c}{$2a\Omega_x=2a\Omega_y=2a\sqrt[4]{\rho h \omega^2/D_{11}}$} \\ 
		\cmidrule(lr){4-10}
		BCs & $\alpha$ & Mode & 1 & 2 & 3 & 4 & 5 & 6 & 7 \\
		\midrule
		SSSS & 0.5 & Mode number  & (1,1) & (1,2) & (1,3) & (1,4) & (1,5) & (1,6) & (1,7) \\
		&     & extended SOV \Citealp{xing2020extended}   & 3.1807 & 3.3190 & 3.5938 & 4.0135 & 4.5495 & 5.1635 & 5.8265 \\
		&     & Present       & 3.1825 & 3.3225 & 3.5975 & 4.0175 & 4.5525 & 5.1625 & 5.8275 \\
		& 1   & Mode number  & (1,1) & (1,2) & (1,3) & (2,1) & (1,4) & (2,2) & (2,3) \\
		&     & extended SOV \Citealp{xing2020extended}   & 3.3190 & 4.0135 & 5.1635 & 6.3615 & 6.5200 & 6.6379 & 7.1876 \\
		&     & Present       & 3.3175 & 4.0175 & 5.1625 & 6.3625 & 6.5175 & 6.6375 & 7.1875 \\
		& 1.5 & Mode number  & (1,1) & (1,2) & (2,1) & (2,2) & (1,3) & (2,3) & (1,4) \\
		&     & extended SOV \Citealp{xing2020extended}   & 3.5938 & 5.1635 & 6.4698 & 7.1876 & 7.2331 & 8.5389 & 9.4352 \\
		&     & Present       & 3.5975 & 5.1675 & 6.4725 & 7.1875 & 7.2325 & 8.5375 & 9.4375 \\
		SCSF & 0.5 & Mode number  & (1,1) & (1,2) & (1,3) & (1,4) & (1,5) & (1,6) & (1,7) \\
		&     & extended SOV \Citealp{xing2020extended}   & 3.1516 & 3.2451 & 3.4588 & 3.8131 & 4.2950 & 4.8711 & 5.5087 \\
		&     & Present      & 3.1525 & 3.2475 & 3.4575 & 3.8175 & 4.2925 & 4.8725 & 5.5075 \\
		& 1   & Mode number  & (1,1) & (1,2) & (1,3) & (1,4) & (2,1) & (2,2) & (2,3) \\
		&     & extended SOV \Citealp{xing2020extended}  & 3.1908 & 3.6428 & 4.5972 & 5.8599 & 6.3033 & 6.4901 & 6.9177 \\
		&     & Present        & 3.1925 & 3.6425 & 4.5975 & 5.8575 & 6.3025 & 6.4925 & 6.9175 \\
		& 1.5 & Mode number & (1,1) & (1,2) & (1,3) & (2,1) & (2,2) & (2,3) & (1,4) \\
		&     & extended SOV \Citealp{xing2020extended}   & 3.2710 & 4.3430 & 6.2157 & 6.3337 & 6.8043 & 7.8718 & 8.3518 \\
		&     & Present      & 3.2725 & 4.3425 & 6.2175 & 6.3325 & 6.8025 & 7.8725 & 8.3525 \\
		GCGC & 0.5 & Mode number & (1,1) & (1,2) & (1,3) & (2,1) & (2,2) & (1,4) & (2,3) \\
		&     & extended SOV \Citealp{xing2020extended}   & 1.1544 & 1.9166 & 2.6835 & 3.1983 & 3.3890 & 3.4501 & 3.7372 \\
		&     & Present      & 1.1525 & 1.9175 & 2.6825 & 3.1975 & 3.3875 & 3.4525 & 3.7375 \\
		& 1   & Mode number  & (1,1) & (2,1) & (1,2) & (2,2) & (1,3) & (2,3) & (3,1) \\
		&     & extended SOV \Citealp{xing2020extended}   & 2.3087 & 3.4900 & 3.8331 & 4.4682 & 5.3669 & 5.7736 & 6.3967 \\
		&     & Present       & 2.3075 & 3.4875 & 3.8325 & 4.4675 & 5.3675 & 5.7725 & 6.3975 \\
		& 1.5 & Mode number  & (1,1) & (2,1) & (1,2) & (2,2) & (3,1) & (3,2) & (1,3) \\
		&     & extended SOV \Citealp{xing2020extended}   & 3.4631 & 4.1353 & 5.7497 & 6.0981 & 6.6049 & 7.6449 & 8.0504 \\
		&     & Present       & 3.4625 & 4.1325 & 5.7475 & 6.0975 & 6.6075 & 7.6425 & 8.0525 \\
		\bottomrule
	\end{tabular}
	\label{tab:sov1}
\end{table}
\begin{table}[!htbp] 
	\centering
	\caption{The first seven frequency parameter $2a\Omega$ of of orthotropic rectangular plates with CCCC, SSCC, SCCC and GGCC boundary conditions.}
	\begin{tabular}{c c l c c c c c c c}
		\toprule
		\multicolumn{3}{c}{} & \multicolumn{7}{c}{$2a\Omega_x=2a\Omega_y=2a\sqrt[4]{\rho h \omega^2/D_{11}}$} \\ 
		\cmidrule(lr){4-10}
		BCs & $\alpha$ & Mode & 1 & 2 & 3 & 4 & 5 & 6 & 7 \\
		\midrule
		CCCC & 0.5 & Mode number  & (1,1) & (1,2) & (1,3) & (1,4) & (1,5) & (1,6) & (1,7) \\
		&     & extended SOV \Citealp{xing2020extended}   & 4.7500 & 4.8208 & 4.9682 & 5.2177 & 5.5791 & 6.0430 & 6.5892 \\
		&     & Present       & 4.7475 & 4.8225 & 4.9725 & 5.2175 & 5.5825 & 6.0425 & 6.5875 \\
		& 1   & Mode number   & (1,1) & (1,2) & (1,3) & (1,4) & (2,1) & (2,2) & (2,3) \\
		&     & extended SOV \Citealp{xing2020extended}   & 4.8579 & 5.3546 & 6.2819 & 7.4972 & 7.9193 & 8.1490 & 8.6054 \\
		&     & Present       & 4.8575 & 5.3575 & 6.2875 & 7.4975 & 7.9175 & 8.1475 & 8.6075 \\
		& 1.5 & Mode number   & (1,1) & (1,2) & (2,1) & (1,3) & (2,2) & (2,3) & (1,4) \\
		&     & extended SOV \Citealp{xing2020extended}   & 5.1581 & 6.5412 & 8.0409 & 8.4945 & 8.7204 & 9.9793 & 10.6460 \\
		&     & Present       & 5.1575 & 6.5375 & 8.0425 & 8.4975 & 8.7175 & 9.9775 & 10.6425 \\
		SSCC & 0.5 & Mode number   & (1,1) & (1,2) & (1,3) & (1,4) & (1,5) & (1,6) & (1,7) \\
		&     & extended SOV \Citealp{xing2020extended}   & 3.9542 & 4.0520 & 4.2525 & 4.5785 & 5.0254 & 5.5682 & 6.1789 \\
		&     & Present       & 3.9575 & 4.0525 & 4.2475 & 4.5775 & 5.0225 & 5.5725 & 6.1825 \\
		& 1   & Mode number  & (1,1) & (1,2) & (1,3) & (1,4) & (2,1) & (2,2) & (2,3) \\
		&     & extended SOV \Citealp{xing2020extended}   & 4.0745 & 4.6606 & 5.7009 & 6.9940 & 7.1396 & 7.3894 & 7.8881 \\
		&     & Present       & 4.0775 & 4.6625 & 5.7025 & 6.9925 & 7.1375 & 7.3875 & 7.8875 \\
		& 1.5 & Mode number   & (1,1) & (1,2) & (2,1) & (1,3) & (2,2) & (2,3) & (1,4) \\
		&     & extended SOV \Citealp{xing2020extended}   & 4.3602 & 5.8384 & 7.2531 & 7.8560 & 7.9481 & 9.2515 & 10.0366 \\
		&     & Present       & 4.3625 & -5.8325 & 7.2525 & 7.8575 & 7.9525 & 9.2525 & 10.0325 \\
		SCCC & 0.5 & Mode number & (1,1) & (1,2) & (1,3) & (1,4) & (1,5) & (1,6) & (1,7) \\
		&     & extended SOV \Citealp{xing2020extended}   & 3.9596 & 4.0745 & 4.3027 & 4.6606 & 5.1361 & 5.7009 & 6.3271 \\
		&     & Present       & 3.9575 & 4.0725 & 4.3025 & 4.6625 & 5.1325 & 5.7025 & 6.3325 \\
		& 1   & Mode number  & (1,1) & (1,2) & (1,3) & (2,1) & (1,4) & (2,2) & (2,3) \\
		&     & extended SOV \Citealp{xing2020extended}   & 4.1349 & 4.8478 & 5.9805 & 7.1541 & 7.3192 & 7.4478 & 8.0121 \\
		&     & Present       & 4.1325 & 4.8475 & 5.9825 & 7.1525 & 7.3175 & 7.4475 & 8.0125 \\
		& 1.5 & Mode number  & (1,1) & (1,2) & (2,1) & (2,2) & (1,3) & (2,3) & (3,1) \\
		&     &extended SOV \Citealp{xing2020extended}  & 4.5824 & 6.2766 & 7.3116 & 8.1528 & 8.3705 & 9.5986 & 10.3507 \\
		&     & Present       & 4.5825 & 6.2775 & 7.3125 & 8.1525 & 8.3725 & 9.5975 & 10.3525 \\
		GGCC & 0.5 & Mode number  & (1,1) & (1,2) & (1,3) & (1,4) & (1,5) & (1,6) & (1,7) \\
		&     & extended SOV \Citealp{xing2020extended}   & 2.3750 & 2.4841 & 2.7895 & 3.2946 & 3.9226 & 4.6123 & 5.3326 \\
		&     & Present       & 2.3725 & 2.4875 & 2.7925 & 3.2975 & 3.9225 & 4.6075 & 5.3325 \\
		& 1   & Mode number  & (1,1) & (1,2) & (1,3) & (2,1) & (2,2) & (1,4) & (2,3) \\
		&     & extended SOV \Citealp{xing2020extended}   & 2.4290 & 3.1410 & 4.4293 & 5.5202 & 5.7315 & 5.8801 & 6.2606 \\
		&     & Present       & 2.4325 & 3.1425 & 4.4325 & 5.5225 & 5.7325 & 5.8775 & 6.2625 \\
		& 1.5 & Mode number  & (1,1) & (1,2) & (2,1) & (2,2) & (1,3) & (2,3) & (3,1) \\
		&     & extended SOV \Citealp{xing2020extended}   & 2.5790 & 4.2472 & 5.5565 & 6.1533 & 6.4347 & 7.5231 & 8.6732 \\
		&     & Present       & 2.5825 & 4.2475 & 5.5575 & 6.1525 & 6.4325 & 7.5225 & 8.6725 \\
		\bottomrule
	\end{tabular}
	\label{tab:sov2}
\end{table}

\begin{table}[!htbp] 
	\centering
	\caption{The first seven nonzero frequency parameter $2a\Omega$ of of orthotropic rectangular plates with CCFF, CFCF, CFFF and FFFF boundary conditions.}
	\begin{tabular}{c c l c c c c c c c}
		\toprule
		\multicolumn{3}{c}{} & \multicolumn{7}{c}{$2a\Omega_x=2a\Omega_y=2a\sqrt[4]{\rho h \omega^2/D_{11}}$} \\ 
		\cmidrule(lr){4-10}
		BCs & $\alpha$ & Mode & 1 & 2 & 3 & 4 & 5 & 6 & 7 \\
		\midrule
		CCFF & 0.5 & Mode number   & (1,1) & (1,2) & (1,3) & (1,4) & (1,5) & (1,6) & (2,1) \\
		&     & extended SOV \Citealp{xing2020extended}   & 1.8978 & 2.0905 & 2.4925 & 3.0563 & 3.7110 & 4.4117 & 4.7029 \\
		&     & Present       & 1.8975 & 2.0925 & 2.4925 & 3.0575 & 3.7125 & 4.4125 & 4.7025 \\
		& 1   & Mode number  & (1,1) & (1,2) & (1,3) & (2,1) & (2,2) & (1,4) & (2,3) \\
		&     & extended SOV \Citealp{xing2020extended}   & 1.9930 & 2.7895 & 4.0733 & 4.7338 & 5.0652 & 5.5128 & 5.7419 \\
		&     & Present       & 1.9925 & 2.7875 & 4.0725 & 4.7325 & 5.0675 & 5.5125 & 5.7425 \\
		& 1.5 & Mode number  & (1,1) & (1,2) & (2,1) & (2,2) & (1,3) & (2,3) & (3,1) \\
		&     & extended SOV \Citealp{xing2020extended}   & 2.1780 & 3.7411 & 4.7931 & 5.5758 & 5.8895 & 7.0263 & 7.9006 \\
		&     & Present       & 2.1775 & 3.7425 & 4.7925 & 5.5725 & 5.8875 & 7.0275 & 7.9025 \\
		CFCF & 0.5 & Mode number   & (1,1) & (1,2) & (1,3) & (1,4) & (1,5) & (1,6) & (1,7) \\
		&     & extended SOV \Citealp{xing2020extended}   & 4.7297 & 4.7427 & 4.7881 & 4.8819 & 5.0478 & 5.3072 & 5.6694 \\
		&     & Present       & 4.7275 & 4.7425 & 4.7875 & 4.8825 & 5.0475 & 5.3075 & 5.6675 \\
		& 1   & Mode number  & (1,1) & (1,2) & (1,3) & (1,4) & (1,5) & (1,6) & (2,1) \\
		&     & extended SOV \Citealp{xing2020extended}   & 4.7295 & 4.7817 & 5.0012 & 5.5348 & 6.4407 & 7.6182 & 7.8523 \\
		&     & Present       & 4.7275 & 4.7825 & 5.0025 & 5.5325 & 6.4425 & 7.6175 & 7.8525 \\
		& 1.5 & Mode number  &(1,1) & (1,2) & (1,3) & (1,4) & (2,1) & (2,2) & (2,3) \\
		&     & extended SOV \Citealp{xing2020extended}   & 4.7292 & 4.8458 & 5.4221 & 6.7635 & 7.8518 & 7.9470 & 8.3021 \\
		&     & Present       & 4.7275 & 4.8475 & 5.4225 & 6.7625 & 7.8525 & 7.9475 & 8.3025 \\
		CFFF & 0.5 & Mode number   & (1,1) & (1,2) & (1,3) & (1,4) & (1,5) & (1,6) & (1,7) \\
		&     & extended SOV \Citealp{xing2020extended}   & 1.8751 & 1.9439 & 2.1679 & 2.5657 & 3.1106 & 3.7486 & 4.4382 \\
		&     & Present       & 1.8775 & 1.9425 & 2.1675 & 2.5675 & 3.1125 & 3.7475 & 4.4375 \\
		& 1   & Mode number  &(1,1) & (1,2) & (1,3) & (1,4) & (2,1) & (2,2) & (2,3) \\
		&     & extended SOV \Citealp{xing2020extended}   & 1.8750 & 2.1242 & 2.9077 & 4.1319 & 4.6937 & 4.8226 & 5.2263 \\
		&     & Present       & 1.8775 & 2.1225 & 2.9075 & 4.1325 & 4.6925 & 4.8225 & 5.2275 \\
		& 1.5 & Mode number  &(1,1) & (1,2) & (1,3) & (2,1) & (2,2) & (2,3) & (1,4) \\
		&     & extended SOV \Citealp{xing2020extended}   & 1.8750 & 2.3402 & 3.8522 & 4.6935 & 4.9753 & 5.8314 & 5.9292 \\
		&     & Present       & 1.8775 & 2.3425 & 3.8525 & 4.6925 & 4.9775 & 5.8325 & 5.9275 \\
		FFFF & 0.5 & Mode number  &(1,3) & (2,2) & (1,4) & (2,3) & (1,5) & (2,4) & (2,5) \\
		&     & extended SOV \Citealp{xing2020extended}   & 1.1540 & 1.4858 & 1.9157 & 2.1704 & 2.6821 & 2.7881 & 3.4093 \\
		&     & Present       & 1.1525 & 1.4875 & 1.9175 & 2.1725 & 2.6825 & 2.7875 & 3.4075 \\
		& 1   & Mode number  &(2,2) & (1,3) & (2,3) & (1,4) & (2,4) & (3,1) & (3,2) \\
		&     & extended SOV \Citealp{xing2020extended}   & 2.1311 & 2.3082 & 3.2734 & 3.8320 & 4.4962 & 4.7298 & 4.9138 \\
		&     & Present       & 2.1325 & 2.3075 & 3.2725 & 3.8325 & 4.4975 & 4.7275 & 4.9125 \\
		& 1.5 & Mode number  &(2,2) & (1,3) & (2,3) & (3,1) & (3,2) & (1,4) & (3,3) \\
		&     & extended SOV \Citealp{xing2020extended}   & 2.6277 & 3.4625 & 4.2915 & 4.7296 & 5.1259 & 5.7485 & 6.1588 \\
		&     & Present       & 2.6275 & 3.4625 & 4.2925 & 4.7275 & 5.1275 & 5.7475 & 6.1575 \\
		\bottomrule
	\end{tabular}
	\label{tab:sov3}
\end{table}
\subsection{Rotationally restrained boundary conditions}
In this subsection, rectangular orthotropic plates with rotationally restrained edges (\(k^v_\xi = k^v_\eta = \infty\)) are validated.  
The rotational stiffness coefficients are defined as: 
% 
\begin{subequations}\label{eq:rotation_coex}
	\begin{align}
		r_{\xi} &= \frac{2a k^r_{\xi}}{D_{11}}, \\
		r_{\eta} &= \frac{2b k^r_{\eta}}{D_{22}}.
	\end{align}
\end{subequations}

The first example considers a square isotropic plate with all four edges rotationally restrained. The vertical translational springs along the four edges are numerically set as \( k^v_{\xi=-1} = k^v_{\xi=1} = k^v_{\eta=-1} = k^v_{\eta=1} = 1 \times 10^{12} \ustif \). The material properties are given as \( D_{11} = D_{22} = D_3 \) and \( v_{12} = v_{21} = 0.3 \).

\Cref{tab:rot1} presents the frequency parameter \( 2a\Omega \) for different rotational stiffness coefficients \( r_{\xi} = r_{\eta} \) with values \( 0.1, 1, 10, 100, \) and \( 1000 \).  
Notably, when \( r_{\xi} = r_{\eta} = 0 \) and \( r_{\xi} = r_{\eta} = \infty \), the boundary conditions correspond to SSSS and CCCC, respectively.

Interestingly, the results indicate that the frequencies \(\Omega_x\) and \(\Omega_y\) are not strictly equal for some mode shapes under rotationally restrained boundary conditions.  
The actual frequency \(\Omega\) lies between \(\Omega_x\) and \(\Omega_y\), which may be attributed to the fact that \(\Omega_x\) and \(\Omega_y\) satisfy Rayleigh’s principle in \Cref{eq:Rayleigh}, representing the weak-form governing equations, but do not necessarily satisfy the strong-form governing equations in \Cref{eq:governing_EOM}.
For a physical problem with exact solutions, both \Cref{eq:Rayleigh,eq:governing_EOM} must be satisfied.  
If this condition is not met, applying \Cref{eq:Rayleigh} still provides a viable approach for approximating the exact solution of the plate.  
Thus, the exact frequency can be estimated as \(\Omega = (\Omega_x + \Omega_y)/2\).  
As shown in \Cref{tab:rot1}, the maximum difference between \(\Omega\) and the solutions reported in \Citealp{zhang2019new} is less than \(1.3\%\).
\Cref{fig:iso} illustrates the variation in mode shapes corresponding to the fundamental natural frequency as the rotational stiffness \(r_{\xi} = r_{\eta}\) increases from zero to \(\infty\), transitioning the boundary conditions from SSSS to CCCC.

\begin{figure}
	\centering
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\textwidth]{figs/iso0.png}
		\caption[]%
		{{\small }}    
		\label{fig:iso0}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.49\textwidth}  
		\centering 
		\includegraphics[width=\textwidth]{figs/iso1.png}
		\caption[]%
		{{\small }}    
		\label{fig:iso1}
	\end{subfigure}
	\vfill
	\begin{subfigure}[b]{0.49\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{figs/iso10.png}
		\caption[]%
		{{\small }}    
		\label{fig:iso10}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.49\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{figs/iso20.png}
		\caption[]%
		{{\small }}    
		\label{fig:iso20}
	\end{subfigure}
	\vskip\baselineskip
	\begin{subfigure}[b]{0.49\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{figs/iso100.png}
		\caption[]%
		{{\small}}    
		\label{fig:iso100}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.49\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{figs/isoinf.png}
		\caption[]%
		{{\small }}    
		\label{fig:isoinf}
	\end{subfigure}
	\caption[]  
	{\small The first mode shape of a square isotropic plate with all four edges rotationally restrained: (a) \(r_{\xi} = r_{\eta} = 0\); (b) \(r_{\xi} = r_{\eta} = 1\); (c) \(r_{\xi} = r_{\eta} = 10\); (d) \(r_{\xi} = r_{\eta} = 20\); (e) \(r_{\xi} = r_{\eta} = 100\); (f) \(r_{\xi} = r_{\eta} = \infty\).}  
	\label{fig:iso}  
\end{figure}

\begin{table}[!htbp]
	\centering
	\caption{The first six frequency parameters, \(2a\Omega = 2a\sqrt[4]{\rho h \omega^2/D_{11}}\), of a square isotropic plate with all four edges rotationally restrained, where \(k^r_{\xi=-1} = k^r_{\xi=1} = k^r_{\eta=-1} = k^r_{\eta=1}\).}  
	\begin{tabular}{c l c c c c c c}
		\toprule
		\multicolumn{1}{c}{} & \multicolumn{6}{c}{$2a\Omega$} \\ 
		\cmidrule(lr){3-8}
		$r_{\xi}=r_{\eta}$ & Mode & 1 & 2 & 3 & 4 & 5 & 6 \\ 
		\midrule
		0.1 & Mode number   & (1,1) & (1,2) & (2,1) & (2,2) & (1,3) & (3,1) \\
		& Ref.\Citealp{mukhopadhyay1979free} & 4.454 & 6.992 & 7.045 & 8.890  & 9.782  & 9.960 \\
		& Ref.\Citealp{zhang2019new} & 4.465 & 7.039 & 7.039 &8.897  &9.945& 9.945 \\
		& Present ($\Omega$)   & 4.463    & 7.035     & 7.035      &8.893     & 9.945     & 9.945   \\
		& Present ($\Omega_x$)   & 4.463     & 7.028     & 7.043     & 8.893     & 9.938       & 9.953  \\
		& Present ($\Omega_y$)   & 4.463    &7.043     &  7.028       & 8.893   &  9.953    & 9.938 \\ 
		&Difference (\%)& 0.044     & 0.056      & 0.056      & 0.044     &  0.000     & 0.000
		\\
		\\
		1 & Mode number   &  (1,1) & (1,2) & (2,1) & (2,2) & (3,1) & (1,3) \\
		& Ref.\Citealp{mukhopadhyay1979free} & 4.529& 7.008 &7.136 &8.936& 9.787& 10.036 \\
		& Ref.\Citealp{zhang2019new} & 4.637 &7.155 & 7.155 & 8.991 & 10.029 & 10.030 \\
		& Present ($\Omega$)   & 4.648      & 7.160   & 7.160     & 8.993    &   10.030   & 10.033   \\
		& Present ($\Omega_x$)   & 4.648     & 7.098      &7.223     & 8.993     &   10.093    & 9.968  \\
		& Present ($\Omega_y$)   & 4.648     & 7.223     & 7.098       & 8.993     &   9.968     & 10.098
		\\ 
		&Difference (\%)& 0.237    & 0.069      & 0.069      & 0.022     &  0.009     &0.029
		\\
		\\
		10 & Mode number   &  (1,1) & (1,2) & (2,1) & (2,2) & (1,3) & (3,1) \\
		& Ref.\Citealp{zhang2019new}  & 5.346 & 7.768 & 7.768 & 9.537 & 10.552 & 10.563 \\
		& Present ($\Omega$)   &5.413      & 7.835     & 7.835     & 9.598   &  10.615    & 10.618  \\
		& Present ($\Omega_x$)   & 5.413      & 7.718      & 7.953     & 9.598     & 10.448      & 10.782 \\
		& Present ($\Omega_y$)   & 5.413     & 7.953     & 7.718      & 9.598    & 10.782      & 10.453
		\\ 
		&Difference (\%)& 1.253     & 0.862      & 0.862      & 0.639     &  0.597     & 0.520  \\
		\\
		100 & Mode number   &  (1,1) & (1,2) & (2,1) & (2,2) & (1,3) & (3,1) \\
		& Ref.\Citealp{mukhopadhyay1979free} & 5.895& 8.326 &8.422 &10.167 &10.957 &11.297 \\
		& Ref.\Citealp{zhang2019new} & 5.901 & 8.442 & 8.442 & 10.253 & 11.307 & 11.333 \\
		& Present ($\Omega$)   & 5.913     & 8.450     & 8.450     & 10.258     &  11.333    & 11.333  \\
		& Present ($\Omega_x$)   & 5.913     & 8.428      & 8.473    & 10.258     & 11.293      & 11.373 \\
		& Present ($\Omega_y$)   & 5.913    & 8.473     & 8.478    & 10.258     & 11.373     & 11.293
		\\ 
		&Difference (\%)& 0.203     & 0.094      & 0.094      & 0.048     & 0.229     & 0.000 \\
		\\
		1000 & Mode number   & (1,1) & (1,2) & (2,1) & (2,2) & (1,3) & (3,1) \\
		& Ref.\Citealp{zhang2019new} & 6.011 & 8.585 & 8.585 & 10.424 & 11.495 & 11.522 \\
		& Present ($\Omega$)   & 5.988      & 8.553    & 8.553     & 10.388    &  11.470     & 11.470  \\
		& Present ($\Omega_x$)   & 5.988     & 8.553      & 8.553      & 10.388      & 11.463      & 11.478 \\
		& Present ($\Omega_y$)   & 5.988     & 8.553     & 8.553    & 10.388      & 11.478       & 11.463
		\\ 
		&Difference (\%)& 0.382    & 0.372      & 0.372      & 0.345     &  0.217     & 0.451
		 \\
		\\
		\bottomrule
	\end{tabular}
	\label{tab:rot1}
\end{table}

The next example considers a rectangular orthotropic plate with three simply supported edges (\(k^r_{\xi=-1} = k^r_{\xi=1} = k^r_{\eta=1} = 0\)), while the edge at \(\eta = -1\) is rotationally restrained. 
The material properties are consistent with those in \citealp{zhang2019new}, where \(2D_{11} = 2D_{22} = D_3\) and \(\nu_{12} = \nu_{21} = 0.3\).  
\Cref{tab:oth2} shows the fundamental frequency results for different length ratios (\(b/a\)), comparing them with those reported in \citealp{zhang2019new}. 
The maximum observed difference is $0.8\%$ when \(r_{\eta=-1} = 10\). 
Furthermore, \Cref{fig:oth} shows how the fundamental mode shape of the square orthotropic plate evolves as the rotational spring stiffness \(k^r_{\eta=-1}\) increases.

In certain numerical calculations involving rotationally restrained boundary conditions, the variables $\alpha_1$ and $\alpha_2$ may take complex values rather than being purely real.  
Consequently, the mode shape coefficients $A_1$, $A_2$, $B_1$, and $B_2$ become complex-valued, leading to $\mathbf{R}$ and $\mathbf{Q}^{-1}$ being complex matrices.  
However, the mode shapes $\phi(\xi)$ and $\psi(\eta)$ remain real-valued, and the dynamic stiffness matrix $\mathbf{K} = \mathbf{R}\mathbf{Q}^{-1}$ is a real symmetric matrix.
Thus, the frequency $\Omega$ can be obtained by solving $\mathbf{K}$ using the refined W-W algorithm provided in this study, which avoids solving the eigenvalue equations in both the real and complex domains.

\begin{table}[!htbp]  
	\centering
	\caption{Fundamental frequency parameter \(2a\Omega = 2a\sqrt[4]{\rho h \omega^2/D_{11}}\) of rectangular orthotropic plates with three edges simply supported (\(k^r_{\xi=-1} = k^r_{\xi=1} = k^r_{\eta=1} = 0\)) and the edge at \(\eta = -1\) rotationally restrained.}
	\begin{tabular}{c c c c c c c }
		\toprule
		\multicolumn{2}{c}{} & \multicolumn{4}{c}{$2a\Omega$} & \\ 
		\cmidrule(lr){3-7}
		$b/a$ & $r_{\eta=-1}$ & Ref.\Citealp{zhang2019new} & Present ($\Omega$) & Present ($\Omega_x$) & Present ($\Omega_y$) & Difference (\%) \\ 
		\midrule
		0.5 & 0         & 7.530 & 7.523 & 7.523 & 7.523 & 0.092 \\ 
		& 1         & 7.690 & 7.700 & 7.588 & 7.813 & 0.130 \\ 
		& 10        & 8.250 & 8.308 & 8.198 & 8.418 & 0.703 \\ 
		& $\infty$  & 8.705 & 8.695 & 8.695 & 8.695 & 0.114 \\ 
		\\
		1.0 & 0         & 4.917 & 4.918 & 4.918 & 4.918 & 0.020 \\ 
		& 1         & 4.954 & 4.960 & 4.933 & 4.988 & 0.121 \\ 
		& 10        & 5.114 & 5.128 & 5.088 & 5.168 & 0.273 \\ 
		& $\infty$  & 5.289 & 5.278 & 5.278 & 5.278 & 0.207 \\ 
		\\
		1.5 & 0         & 4.126 & 4.128 & 4.128 & 4.128 & 0.048 \\ 
		& 1         & 4.139 & 4.138 & 4.128 & 4.148 & 0.024 \\ 
		& 10        & 4.202 & 4.208 & 4.188 & 4.228 & 0.142 \\ 
		& $\infty$  & 4.292 & 4.288 & 4.288 & 4.288 & 0.093 \\ 
		\bottomrule
	\end{tabular}
	\label{tab:oth2}
\end{table}
\begin{figure}
	\centering
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\textwidth]{figs/otho0.png}
		\caption[]%
		{{\small }}    
		\label{fig:otho0}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.49\textwidth}  
		\centering 
		\includegraphics[width=\textwidth]{figs/otho1.png}
		\caption[]%
		{{\small }}    
		\label{fig:otho1}
	\end{subfigure}
	\vfill
	\begin{subfigure}[b]{0.49\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{figs/otho10.png}
		\caption[]%
		{{\small }}    
		\label{fig:otho10}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.49\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{figs/otho20.png}
		\caption[]%
		{{\small }}    
		\label{fig:otho20}
	\end{subfigure}
	\vskip\baselineskip
	\begin{subfigure}[b]{0.49\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{figs/otho100.png}
		\caption[]%
		{{\small}}    
		\label{fig:otho100}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.49\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{figs/othoinf.png}
		\caption[]%
		{{\small }}    
		\label{fig:othoinf}
	\end{subfigure}
	\caption[]  
	{\small The first mode shape of a square orthotropic plate with the edge at \(\eta=-1\) rotationally restrained: (a) \(r_{\eta=-1} = 0\); (b) \(r_{\eta=-1} = 1\); (c) \(r_{\eta=-1} = 10\); (d) \(r_{\eta=-1} = 20\); (e) \(r_{\eta=-1} = 100\); (f) \(r_{\eta=-1} = \infty\).}  
	\label{fig:oth}  
\end{figure}
\FloatBarrier
\section{Conclusion}
In this study, the dynamic stiffness matrix (DSM) based on the extended separation-of-variable (SOV) mode functions has been developed for the vibration analysis of an orthotropic rectangular plate with general homogeneous boundary conditions.

Based on the Rayleigh quotient, the governing equation of the plate, along with the shear force and bending moment equilibrium related to the boundary conditions, are derived.  
The extend-SOV type solutions obtained from the governing equation are then used to develop the dynamic stiffness matrices by applying the boundary conditions.
Instead of solving eigenvalue equations involving highly nonlinear and transcendental functions derived from the determinant of the dynamic stiffness matrix, a refined W-W algorithm based on the 'hypothetical structure method' is developed to address this eigenvalue problem.
The challenge of determining the fully clamped frequencies using the W-W algorithm is resolved by finding the simply supported frequencies, whose closed-form expression can be easily derived based on the SOV method.  
Due to the improved W-W algorithm and the concise SOV-form solutions, the process of solving the dynamic stiffness matrix becomes more systematic and efficient.

Classical boundary conditions, such as guided, simply supported, clamped, and free edges, can be realized by setting the translational springs (\(k^v_\xi\)) and rotational springs (\(k^r_\xi\)) along the plate edges to either zero or infinity.  
Numerical results confirm that accurate solutions are obtained for these boundary conditions.  
Despite some approximations in certain elastically restrained boundary conditions, the maximum percentage error across all numerical experiments remains within $1.25\%$.  
This may occur because the SOV-form solutions used are derived from the weak-form governing equation, which is based on Rayleigh's principle.

Since the SOV-form solution \(\phi(\xi) \psi(\eta)\) consists of only a single term, unlike the infinite series expansions used in the traditional DSM, each eigenvalue solution can be explicitly expressed.  
This suggests the potential for obtaining concise and closed-form solutions for assembled plate structures compared to traditional DSM methods.

\FloatBarrier
\begin{appendices}
\numberwithin{equation}{section}
\renewcommand{\theequation}{\thesection.\arabic{equation}}
\section{Integral parameters}\label{appendA}
The integral parameters \( S_1, S_2, S_3, \) and \( S_4 \) are defined as follows:
%
\begin{equation}\label{eq:inte_phi}
	\begin{aligned}
		S_1 &= \int_{0}^{1} \psi^2 \, d\eta \\
		&= (B_1^2 + B_2^2 - B_3^2 + B_4^2) 
		+ \frac{-B_1^2 + B_2^2}{2\alpha_2} \sin(2\alpha_2) 
		+ \frac{B_3^2 + B_4^2}{2\beta_2} \sinh(2\beta_2) \\
		&\quad + \frac{4(\alpha_2 B_2B_4 + \beta_2 B_1B_3)}{\alpha_2^2 + \beta_2^2} \sin(\alpha_2) \cosh(\beta_2) \\
		&\quad + \frac{4(-\alpha_2 B_1B_3 + \beta_2 B_2B_4)}{\alpha_2^2 + \beta_2^2} \cos(\alpha_2) \sinh(\beta_2).
	\end{aligned}
\end{equation}
%
\begin{equation}\label{eq:inte_dphi2}
\begin{split}
	S_2 &= \int_{0}^{1} \left( \psi \frac{d^2 \psi}{d \eta^2} \right) d \eta \\
	&= \left( -\alpha_2^2 B_1^2 - \alpha_2^2 B_2^2 - \beta_2^2 B_3^2 + \beta_2^2 B_4^2 \right) \\
	&\quad + \frac{\alpha_2 (B_1^2 - B_2^2)}{2} \sin(2 \alpha_2) 
	 + \frac{\beta_2 (B_3^2 + B_4^2)}{2} \sinh(2 \beta_2) \\
	&\quad + \frac{2(-\alpha_2^2 + \beta_2^2) (\alpha_2 B_2 B_4 + \beta_2 B_1 B_3)}{\alpha_2^2 + \beta_2^2} \sin(\alpha_2) \cosh(\beta_2) \\
	&\quad + \frac{2(-\alpha_2^2 + \beta_2^2) (-\alpha_2 B_1 B_3 + \beta_2 B_2 B_4)}{\alpha_2^2 + \beta_2^2} \cos(\alpha_2) \sinh(\beta_2).
\end{split}
\end{equation}
%
\begin{equation}\label{eq:inte_dphi3}
	\begin{aligned}
		S_3 &= \int_{0}^{1} \left(\frac{d\psi}{d\eta}\right)^2 \, d\eta \\
		&= \alpha_2^2 B_1^2 + \alpha_2^2 B_2^2 + \beta_2^2 B_3^2 - \beta_2^2 B_4^2 \\
		&\quad + \frac{\alpha_2 (B_1^2 - B_2^2)}{2} \sin(2\alpha_2) 
		 + \frac{\beta_2 (B_3^2 + B_4^2)}{2} \sinh(2\beta_2) \\
		&\quad + \frac{4 \alpha_2 \beta_2 (\alpha_2 B_1 B_3 - \beta_2 B_2 B_4)}{\alpha_2^2 + \beta_2^2} \sin(\alpha_2) \cosh(\beta_2) \\
		&\quad + \frac{4 \alpha_2 \beta_2 (\alpha_2 B_2 B_4 + \beta_2 B_1 B_3)}{\alpha_2^2 + \beta_2^2} \cos(\alpha_2) \sinh(\beta_2).
	\end{aligned}
\end{equation}
%
\begin{equation}\label{eq:inte_dphi4}
	\begin{split}
		S_4 &= \int_{0}^{1} \left( \frac{d^2 \psi}{d \eta^2} \right)^2 d \eta \\
		&= \left( \alpha_2^4 B_1^2 + \alpha_2^4 B_2^2 - \beta_2^4 B_3^2 + \beta_2^4 B_4^2 \right) \\
		&\quad + \frac{\alpha_2^3 (-B_1^2 + B_2^2)}{2} \sin(2 \alpha_2)
		+ \frac{\beta_2^3 (B_3^2 + B_4^2)}{2} \sinh(2 \beta_2) \\
		&\quad + \frac{4 \alpha_2^2 \beta_2^2 (-\alpha_2 B_2 B_4 - \beta_2 B_1 B_3)}{\alpha_2^2 + \beta_2^2} \sin(\alpha_2) \cosh(\beta_2) \\
		&\quad + \frac{4 \alpha_2^2 \beta_2^2 (\alpha_2 B_1 B_3 - \beta_2 B_2 B_4)}{\alpha_2^2 + \beta_2^2} \cos(\alpha_2) \sinh(\beta_2)
	\end{split}
\end{equation}
%
The integral parameters \( T_1 \), \( T_2 \), \( T_3 \), and \( T_4 \) can be obtained by replacing \( B_1 \) to \( B_4 \) by \( A_1 \) to \( A_4 \), respectively, and \( \alpha_2 \) and \( \beta_2 \) by \( \alpha_1 \) and \( \beta_1 \), respectively.

\end{appendices}
\FloatBarrier
\bibliographystyle{elsarticle-harv}
\bibliography{bibliography}

\end{document}