### 2025-03-31 figs/general_plate_tikz.tex

\begin{tikzpicture}[yscale=1]
	\tikzset{
	pics/boundary-box/.style={
		code={
			\begin{scope}[shift={(0mm,0mm)}]
				\definecolor{fc}{RGB}{204,204,204}
				\draw [fill=none, draw=black, line width=0.6mm] (000.0mm,0.0mm)
				-- ++(0.0mm,60.0mm)
				-- ++(-60.0mm,0.0mm)
				-- ++(0.0mm,-60.0mm)  % Adjusted the vertical length to 20mm
				-- ++(60.0mm,0.0mm)  % Adjusted the vertical length to 20mm
				-- cycle;
				\node [black, anchor=west, font=\fontsize{100}{0}\selectfont\bfseries] at (-40mm, 30mm) { $\textbf{?}$ };
			\end{scope}
		}
	}
}
\tikzset{
	pics/Mass-circle/.style={
		code={
			\draw[fill=gray!30, draw=black, line width=0.6mm] (0mm, 0mm) circle [radius=38mm];
		}
	}
}
\tikzset{
	pics/Beam/.style={
		code={\draw [fill=none,draw=black,line width=10.0mm] (000.0mm,0.mm)
			-- ++(600.0mm,0.0mm)
			;
			%% curvature
			\draw[fill=none, draw=gray, line width=4.0mm, dash pattern=on 20pt off 20pt] (000.0mm,0.mm) .. controls (200mm,-100mm) and (300mm,-50mm) .. (600.0mm,0mm);

		}
	}
}
\tikzset{
	pics/Spring-damp/.style={
		code={\begin{scope}[cm={0.4,0,0,0.4,(40.0mm,0mm)}]
				\draw [fill=none,draw=black,line width=2.0mm] (145.0mm,707.4mm)
				-- ++(0.0mm,33.0mm)
				;
				\draw [fill=none,draw=black,line width=2.0mm] (120.0mm,867.4mm)
				-- ++(50.0mm,-15.0mm)
				;
				\draw [fill=none,draw=black,line width=2.0mm] (170.0mm,852.4mm)
				-- ++(-50.0mm,-15.0mm)
				;
				\draw [fill=none,draw=black,line width=2.0mm] (120.0mm,837.4mm)
				-- ++(50.0mm,-15.0mm)
				;
				\draw [fill=none,draw=black,line width=2.0mm] (170.0mm,822.4mm)
				-- ++(-50.0mm,-15.0mm)
				;

				\draw [fill=none,draw=black,line width=2.0mm] (120.0mm,807.4mm)
				-- ++(50.0mm,-15.0mm)
				;

				\draw [fill=none,draw=black,line width=2.0mm] (170.0mm,792.4mm)
				-- (120.0mm,777.4mm)
				;

				\draw [fill=none,draw=black,line width=2.0mm] (120.0mm,777.4mm)
				-- ++(50.0mm,-15.0mm)
				;

				\draw [fill=none,draw=black,line width=2.0mm] (170.0mm,762.4mm)
				-- ++(-50.0mm,-15.0mm)
				;

				\draw [fill=none,draw=black,line width=2.0mm] (145.0mm,875.4mm)
				-- ++(-25.0mm,-8.0mm)
				;

				\draw [fill=none,draw=black,line width=2.0mm] (120.0mm,747.4mm)
				-- ++(25.0mm,-7.0mm)
				;

				\draw [fill=none,draw=black,line width=2.0mm] (145.0mm,875.4mm)
				-- (145.0mm,907.4mm)
				;
			\end{scope}
			% horizontal line above mass left
			\draw [fill=none,draw=black,line width=2.0mm] (58.0mm,363mm)
			-- ++(40.0mm,0.0mm)
			;

			%% damping left
			\begin{scope}[cm={0.4,0,0,-0.4,(0.0mm,645.9mm)}]
				\draw [fill=none,draw=black,line width=2.0mm] (115.0mm,827.7mm)
				-- ++(0.0mm,-35.3mm)
				-- (175.0mm,792.4mm)
				-- ++(0.0mm,35.0mm)
				;

				\draw [fill=none,draw=black,line width=2.0mm] (145.0mm,707.4mm)
				-- ++(0.0mm,85.0mm)
				;

				\draw [fill=none,draw=black,line width=2.0mm] (123.0mm,807.4mm)
				-- ++(44.0mm,0.0mm)
				;

				\draw [fill=none,draw=black,line width=2.0mm] (145.0mm,807.4mm)
				-- ++(0.0mm,100.0mm)
				;
			\end{scope}

			%% suspension arrow left
			\draw [fill=none,draw=black,line width=2.0mm] (58.0mm,283.4mm)
			-- ++(40.0mm,0.0mm)
			;
			\draw [fill=none,draw=black,line width=2.0mm] (78.0mm,363.4mm)
			-- ++(0mm,27.0mm)
			;
			%% vertical line upp
			\draw [fill=none,draw=black,line width=2.0mm] (78.0mm,283.4mm)
			-- ++(0.0mm,-30.0mm)
			;
		}
	}
}
\tikzset{
	pics/beam-Spring/.style={
		code={\begin{scope}[cm={0.4,0,0,0.4,(-38.0mm,-256.4mm)}]
				\draw [fill=none,draw=black,line width=2.0mm] (145.0mm,782.4mm)
				-- ++(0.0mm,-60.0mm)
				;
				\draw [fill=none,draw=black,line width=2.0mm] (120.0mm,867.4mm)
				-- ++(50.0mm,-15.0mm)
				;
				\draw [fill=none,draw=black,line width=2.0mm] (170.0mm,852.4mm)
				-- ++(-50.0mm,-15.0mm)
				;
				\draw [fill=none,draw=black,line width=2.0mm] (120.0mm,837.4mm)
				-- ++(50.0mm,-15.0mm)
				;
				\draw [fill=none,draw=black,line width=2.0mm] (170.0mm,822.4mm)
				-- ++(-50.0mm,-15.0mm)
				;
				\draw [fill=none,draw=black,line width=2.0mm] (120.0mm,807.4mm)
				-- ++(50.0mm,-15.0mm)
				;
				\draw [fill=none,draw=black,line width=2.0mm] (170.0mm,792.4mm)
				-- (145.0mm,782.4mm)
				;

				\draw [fill=none,draw=black,line width=2.0mm] (145.0mm,875.4mm)
				-- ++(-25.0mm,-8.0mm)
				;

				\draw [fill=none,draw=black,line width=2.0mm] (145.0mm,875.4mm)
				-- (145.0mm,950.4mm)
				;
			\end{scope}

		}
	}
}
\tikzset{
	pics/beam-arrow/.style={
		code={% bridge curve property 
			\draw [fill=none,draw=black,line width=2.0mm] (380.0mm,32.0mm)
			.. controls ++(85.0mm,-8.0mm) and ++(-60.0mm,-6.0mm) .. ++(135.0mm,-32.0mm)
			;
			%% arraw of bridge curve property 
			\draw [->, black, line width=6mm, >=stealth, shorten >=1pt] (515.0mm,0mm) -- ++(0.5,0); % Draws the arrowhead separately

		}
	}
}
\tikzset{
	pics/beam-prop-arrow/.style={
		code={% bridge curve deflection yx
			\draw [fill=none,draw=black,line width=2.0mm] (440.0mm,0.0mm)
			.. controls ++(85.0mm,5.0mm) and ++(-60.0mm,4.0mm) .. ++(135.0mm,36.0mm)
			;
			%% arraw of bridge curve deflection yx
			\draw [black, line width=2mm] (570.0mm,36.0mm) -- (580.0mm,36.0mm);
			\draw [->, black, line width=6mm, >=stealth, shorten >=1pt] (580.0mm,36.0mm) -- ++(0.5,0); % Draws the arrowhead separately

		}
	}
}
\tikzset{
	pics/mass-defle/.style={
		code={%% arraw of vehicle deflection yv 
			\draw [->,fill=none,draw=black,line width=2.50mm] (0.0mm,25.0mm)
			-- (30.0mm,25.0mm)
			-- (30.0mm,65.3mm)
			;			
		}
	}
}
\tikzset{
	pics/floor/.style={
		code={
			%  left
			\begin{scope}[cm={0.4,0,0,-0.4,(-42.0mm,377.0mm)}]
				\draw [fill=none,draw=black,line width=2.0mm] (30.0mm,939mm)
				-- (190.0mm,939mm)
				;
				\draw [fill=none,draw=black,line width=2.0mm] (50.0mm,939mm)
				-- (30.0mm,965mm)
				;
				\draw [fill=none,draw=black,line width=2.0mm] (80.0mm,939mm)
				-- (60.0mm,965mm)
				;
				\draw [fill=none,draw=black,line width=2.0mm] (110.0mm,939mm)
				-- (90.0mm,965mm)
				;
				\draw [fill=none,draw=black,line width=2.0mm] (140.0mm,939mm)
				-- (120.0mm,965mm)
				;
				\draw [fill=none,draw=black,line width=2.0mm] (170.0mm,939mm)
				-- (150.0mm,965mm)
				;
			\end{scope}

		}
	}
}
\tikzset{
	pics/beam-rotation/.style={
		code={
			%%% arc representing rotation or deflection
			\draw [-,fill=none,draw=black,line width=2.00mm] 
			(0.0mm,25.0mm) arc [start angle=270, end angle=90, radius=30mm];
			\draw [fill=none,draw=black,line width=2.5mm] (0.0mm,60mm)
			-- (0.0mm,100mm)
			;

			\draw [fill=none,draw=black,line width=2.5mm] (0.0mm,70mm)
			-- (15.0mm,75mm)
			;
			\draw [fill=none,draw=black,line width=2.5mm] (0.0mm,80mm)
			-- (15.0mm,85mm)
			;
			\draw [fill=none,draw=black,line width=2.5mm] (0.0mm,90mm)
			-- (15.0mm,95mm)
			;

		}
	}
}
\tikzset{
	pics/beam-rotation-bot/.style={
		code={
			%%% arc representing rotation or deflection
			\draw [-,fill=none,draw=black,line width=2.00mm] 
			(0.0mm,25.0mm) arc [start angle=270, end angle=50, radius=30mm];
			\draw [fill=none,draw=black,line width=2.5mm] (0.0mm,60mm)
			-- (40.0mm,100mm)
			;

			\draw [fill=none,draw=black,line width=2.5mm] (10mm,70mm)
			-- (20.0mm,60mm)
			;
			\draw [fill=none,draw=black,line width=2.5mm] (20mm,80mm)
			-- (30.0mm,70mm)
			;
			\draw [fill=none,draw=black,line width=2.5mm] (30mm,90mm)
			-- (40.0mm,80mm)
			;

		}
	}
}
\tikzset{
	pics/Rectangular-Plate/.style={
		code={
			% Define the vertices of the rectangular plate
			\coordinate (A) at (0, 1, 0);
			\coordinate (B) at (90, 1, 0);
			\coordinate (C) at (90, 2, 0);
			\coordinate (D) at (0, 2, 0);
			\coordinate (A') at (0, 1, 80);
			\coordinate (B') at (90, 1, 80);
			\coordinate (C') at (90, 2, 80);
			\coordinate (D') at (0, 2, 80);
			\coordinate (E) at (0, 0, 0);
			\coordinate (F) at (90, 0, 0);
			\coordinate (E') at (0, 0, 80);
			\coordinate (F') at (90, 0, 80);

			% Fill the surface AA'B'B with gray and transparent
			\fill[gray!50, fill opacity=0.2] (A) -- (A') -- (B') -- (B) -- cycle;

			% Draw the edges of the bottom face
			\draw[black, line width=2mm, dashed, dash pattern=on 10mm off 5mm] (A) -- (B); % AB dashed
			\draw[black, line width=2mm] (B) -- (C) -- (D); % Remaining edges solid
			\draw[black, line width=2mm, dashed, dash pattern=on 10mm off 5mm] (D) -- (A); % AD dashed
			\draw[black, line width=2mm, dashed, dash pattern=on 10mm off 5mm] (A) -- (E); % AB dashed
			\draw[black, line width=2mm, dashed, dash pattern=on 10mm off 5mm] (F) -- (E); % AB dashed
			\draw[black, line width=2mm, dashed, dash pattern=on 10mm off 5mm] (F) -- (B); % AB dashed
			\draw[black, line width=2mm, dashed, dash pattern=on 10mm off 5mm] (E') -- (E); % AB dashed

			% Draw the edges of the top face
			\draw[black, line width=2mm] (A') -- (B') -- (C') -- (D') -- cycle;

			% Draw the vertical edges connecting the bottom and top faces
			\draw[black, line width=2mm, dashed, dash pattern=on 10mm off 5mm] (A) -- (A'); % AA' dashed
			\draw[black, line width=2mm] (B) -- (B');
			\draw[black, line width=2mm] (C) -- (C');
			\draw[black, line width=2mm] (D) -- (D');
			\draw[black, line width=2mm] (F) -- (F');
			\draw[black, line width=2mm] (A') -- (E');
			\draw[black, line width=2mm] (B') -- (F');
			\draw[black, line width=2mm] (F') -- (E');
		}
	}
}

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% drawing
	%\pic at (700mm,0mm) {boundary-box};
	%\pic at (40mm,0mm) {boundary-box};
	\pic at (20mm,-25mm) {beam-rotation};
	\pic at (0mm,-120mm) {beam-Spring};
	\pic at (20mm,-90mm) {floor};
	\pic at (-60mm,-105mm) {beam-rotation};
	\pic at (-80mm,-200mm) {beam-Spring};
	\pic at (-60mm,-170mm) {floor};
	\pic at (-140mm,-185mm) {beam-rotation};
	\pic at (-160mm,-280mm) {beam-Spring};
	\pic at (-140mm,-250mm) {floor};

	%%%%%%%%%%%%
	\pic[xscale=-1] at (900mm,-35mm) {beam-rotation};
	\pic at (880mm,-130mm) {beam-Spring};
	\pic at (900mm,-100mm) {floor};
	\pic[xscale=-1] at (820mm,-115mm) {beam-rotation};
	\pic at (810mm,-210mm) {beam-Spring};
	\pic at (820mm,-180mm) {floor};
	\pic[xscale=-1] at (730mm,-205mm) {beam-rotation};
	\pic at (720mm,-300mm) {beam-Spring};
	\pic at (730mm,-270mm) {floor};
%%%%%%%%%%%%%%%%%%%%%%%%
	\pic[xscale=-1] at (160mm,35mm) {beam-rotation-bot};
	\pic at (140mm,-65mm) {beam-Spring};
	\pic at (160mm,-35mm) {floor};
	\pic[xscale=-1] at (360mm,35mm) {beam-rotation-bot};
	\pic at (340mm,-65mm) {beam-Spring};
	\pic at (360mm,-35mm) {floor};
	\pic[xscale=-1] at (560mm,35mm) {beam-rotation-bot};
	\pic at (540mm,-65mm) {beam-Spring};
	\pic at (560mm,-35mm) {floor};
	\pic[xscale=-1] at (760mm,35mm) {beam-rotation-bot};
	\pic at (740mm,-65mm) {beam-Spring};
	\pic at (760mm,-35mm) {floor};
	%%%%%%%%%
	\pic[xscale=1] at (510mm,-275mm) {beam-rotation-bot};
	\pic at (490mm,-375mm) {beam-Spring};
	\pic at (510mm,-345mm) {floor};
	\pic[xscale=1] at (310mm,-275mm) {beam-rotation-bot};
	\pic at (290mm,-375mm) {beam-Spring};
	\pic at (310mm,-345mm) {floor};
	\pic[xscale=1] at (110mm,-275mm) {beam-rotation-bot};
	\pic at (90mm,-375mm) {beam-Spring};
	\pic at (110mm,-345mm) {floor};
	\pic[xscale=1] at (-90mm,-275mm) {beam-rotation-bot};
	\pic at (-110mm,-375mm) {beam-Spring};
	\pic at (-90mm,-345mm) {floor};

	%%%%%%%%
    \pic at (3mm, 6, 3) {Rectangular-Plate};

    %%%%%%%%%%%%%%%%%%%%%
    %%% 2b
    \begin{scope}[shift={(-35mm,-180.0mm)}]
    	\draw[{Latex[length=18mm, width=12mm]}-{Latex[length=18mm, width=12mm]}, fill=none, draw=black, line width=2.5mm] 
    	(700mm,-270.0mm) -- (1010mm,60.0mm);
    	\draw [-,fill=none,draw=black,line width=2.mm] (700mm,-150.0mm)
    	-- (700mm,-300.0mm)
    	;
    	\draw [-,fill=none,draw=black,line width=2.mm] (1010mm,90.0mm)
    	-- (1010mm,000.0mm)
    	;
    \end{scope}
    %%% 2a
    \begin{scope}[shift={(-35mm,150.0mm)}]
    	\draw[{Latex[length=18mm, width=12mm]}-{Latex[length=18mm, width=12mm]}, fill=none, draw=black, line width=2.5mm] 
    	 (110mm,50.0mm)
    	-- (1010mm,50.0mm);
    	\draw [-,fill=none,draw=black,line width=2.mm] (110mm,70.0mm)
    	-- (110mm,-30.0mm)
    	;
    	\draw [-,fill=none,draw=black,line width=2.mm] (1010mm,70.0mm)
    	-- (1010mm,-30.0mm)
    	;
    \end{scope}
        %%% h
    \begin{scope}[shift={(-35mm,150.0mm)}]
    	\draw[{Latex[length=12mm, width=8mm]}-{Latex[length=12mm, width=8mm]}, fill=none, draw=black, line width=2.5mm] 
    	(-250mm,-385.0mm)
    	-- (-250mm,-415.0mm);
    	\draw [-,fill=none,draw=black,line width=2.mm] (-260mm,-388.0mm)
    	-- (-200mm,-388.0mm)
    	;
    	\draw [-,fill=none,draw=black,line width=2.mm] (-260mm,-410.0mm)
    	-- (-200mm,-410.0mm)
    	;
    \end{scope}
     %%% xyz
    \begin{scope}[shift={(30mm,-150.0mm)}]
    	\draw[-{Latex[length=18mm, width=12mm]}, fill=none, draw=black, line width=2.5mm, dashed, dash pattern=on 10mm off 5mm] 
    	(360mm,50.0mm) -- (460mm,50.0mm);
    	\draw[-{Latex[length=18mm, width=12mm]}, fill=none, draw=black, line width=2.5mm] 
    	(360mm, 50.0mm) -- (360mm, 150.0mm); 		
    	\draw[-{Latex[length=18mm, width=12mm]}, fill=none, draw=black, line width=2.5mm, dashed, dash pattern=on 10mm off 5mm] 
    	(360mm, 50.0mm) -- (420mm, 110.0mm);

    \end{scope}
	%%%%%%%%%%%%%%%% text

	\begin{scope}[shift={(16.9mm,-8.4mm)}]
		\node [black,anchor=west,font=\fontsize{140}{0}\selectfont] at (340mm,-110mm) { $o$ };
		\node [black,anchor=west,font=\fontsize{140}{0}\selectfont] at (460mm,-120mm) { $x$ };
		\node [black,anchor=west,font=\fontsize{140}{0}\selectfont] at (430mm,-20mm) { $y$ };
		\node [black,anchor=west,font=\fontsize{140}{0}\selectfont] at (380mm,30mm) { $z$ };
		\node [black,anchor=west,font=\fontsize{140}{0}\selectfont] at (460mm,240mm) { $2a$ };
		\node [black,anchor=west,font=\fontsize{140}{0}\selectfont] at (860mm,-250mm) { $2b$ };
		\node [black,anchor=west,font=\fontsize{140}{0}\selectfont] at (-350mm,-240mm) { $h$ };
		\node [black,anchor=west,font=\fontsize{140}{0}\selectfont] at (-50mm,-140mm) { $k^v_{\xi=-1}$ };
		\node [black,anchor=west,font=\fontsize{140}{0}\selectfont] at (-200mm,20mm) { $k^r_{\xi=-1}$ };
		\node [black,anchor=west,font=\fontsize{140}{0}\selectfont] at (750mm,-220mm) { $k^v_{\xi=1}$ };
	    \node [black,anchor=west,font=\fontsize{140}{0}\selectfont] at (640mm,-50mm) { $k^r_{\xi=1}$ };
	    \node [black,anchor=west,font=\fontsize{140}{0}\selectfont] at (320mm,-300mm) { $k^v_{\eta=-1}$ };
		\node [black,anchor=west,font=\fontsize{140}{0}\selectfont] at (200mm,-150mm) { $k^r_{\eta=-1}$ };
		\node [black,anchor=west,font=\fontsize{140}{0}\selectfont] at (170mm,0mm) { $k^v_{\eta=1}$ };
		\node [black,anchor=west,font=\fontsize{140}{0}\selectfont] at (160mm,150mm) { $k^r_{\eta=1}$ };
	\end{scope}
\end{tikzpicture}